---
title: "Differential expression analysis Big Clusters"
author: "Camila Arcuschin; Ignacio Schor"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    theme: cerulean
    code_folding: hide
out.width: '30%'
out.heigh: '100%'
fig.width: '100%'
fig.height: '100%'
dpi: 100
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = TRUE, warning = FALSE)
knitr::opts_knit$set(root.dir = "/data_husihuilke/shared/SMB/carcu/Parkinsons_project/")
```

```{r, include=FALSE}
library(Seurat)
library(tidyverse)
library(DESeq2)
library(zinbwave)
library(scran)
library(RhpcBLASctl)
library(ggrepel)
library(BiocParallel)
library(DT)
register(MulticoreParam(10))
```

```{r, include=FALSE}
#### Functions
FCfun <- function (res_DE, fc, pad){
  res_DE$diffexpressed <- "NO"
  res_DE$diffexpressed[res_DE$log2FC > fc & res_DE$padj <= pad] <- "UP"
  res_DE$diffexpressed[res_DE$log2FC < -fc & res_DE$padj <= pad] <- "DOWN"
  res_DE$delabel <- NA
  res_DE$delabel[res_DE$diffexpressed != "NO"] <- res_DE$Gene[res_DE$diffexpressed != "NO"]
  res_DE$delabel[res_DE$Gene == "hSNCA-A53T"] <- "hSNCA-A53T"

  Volcano<-ggplot(data=res_DE , aes(x=log2FC, y=-log10(padj), col=diffexpressed, label=delabel)) +
    geom_point() +
    theme_minimal() +
    #xlim(min, max) +
    geom_text_repel(max.overlaps = 100) +
    scale_color_manual(values=c("DOWN" = "blue", "NO" = "black", "UP" = "red")) +
    geom_vline(xintercept=c(-fc, fc), col="red") +
    geom_hline(yintercept=-log10(pad), col="red")
  Volcano
}

## Exclude unwellcomed genes from labeling, in the effect measure
FCfun2 <- function (res_DE, fc, pad, notWellcome){
  res_DE$diffexpressed <- "NO"
  res_DE$diffexpressed[res_DE$log2FC > fc & res_DE$padj <= pad] <- "UP"
  res_DE$diffexpressed[res_DE$log2FC < -fc & res_DE$padj <= pad] <- "DOWN"
  res_DE$delabel <- NA
  res_DE$delabel[res_DE$diffexpressed != "NO"] <- res_DE$Gene[res_DE$diffexpressed != "NO"]
  res_DE$delabel[res_DE$Gene %in% notWellcome] <- NA
  
  Volcano<-ggplot(data=res_DE, aes(x=log2FC, y=-log10(padj), col=diffexpressed, label=delabel, alpha = !(Gene %in% notWellcome))) +
    geom_point() +
    theme_minimal() +
    #xlim(min, max) +
    geom_text_repel(max.overlaps = 100) + 
    scale_color_manual(values=c("DOWN" = "blue", "NO" = "black", "UP" = "red")) +
    geom_vline(xintercept=c(-fc, fc), col="red") +
    geom_hline(yintercept=-log10(pad), col="red")
  Volcano
}
```

## Subseted data for post-hoc analysis of the jmo

Differentially expressed genes (padj <= 0.05) from the analysis of the jmo, clust m, were used to check differences in the medias only across treatment (no interaction with DS evaluated).

```{r}
m = args[1]
j = args[2]
file = args[3]
date = args[4]
DESeq2_output <- read_csv(paste0("Outputs/", date, "_FCmat_official_full_log2FC_clust", m, "_", j, "_LC.csv"))
nonDE_genes <- filter(DESeq2_output, padj > 0.05 | is.na(padj))
nonDE_genes <- nonDE_genes$Gene
```


## DE Analysis

[The model:]{.underline}

For DE Analysis, we are doing an nbinomLRT. This means that we are doing
a Likelihood Ratio Test of a full vs. a reduced model that both of them
fit the data to a negative binomial model.

In this context, the full model is:
$$Counts  ⁓   Treat$$
And the reduced model is: $$Counts  ⁓  1$$


Same control of droupouts method was used as before (analysis from the 230626), also this counts for parameters used to estimate dispersion


```{r CoreAnalysis, message=FALSE}

  SeuObj <- readRDS(paste0("Outputs/", file))
  DefaultAssay(SeuObj) <- "RNA"
  SeuObj <- SetAssayData(SeuObj, assay.type = "RNA", slot = "counts", new.data = round(GetAssayData(SeuObj, slot = "counts"), 0))
  #GetAssayData(object = SeuObj, slot = "counts")
  SeuObj@meta.data <- select(SeuObj@meta.data, seurat_clusters, Treat, Age, orig.ident)
  genenames <- rownames(SeuObj[["RNA"]]) 
  SceLC <- subset(SeuObj, subset = seurat_clusters == m & Age == j)
  SceLC <- DietSeurat(SceLC)
  SceLC <- as.SingleCellExperiment(SceLC)
  Average_Counts <- rowMeans(assays(SceLC)[["counts"]]) %>% log10()
  ExpressingCells <- rowSums(assays(SceLC)[["counts"]]>0)
  PropExpressingCells <- ExpressingCells/dim(SceLC)[2]
    if (Average_Counts["hSNCA-A53T"] > -1 & PropExpressingCells["hSNCA-A53T"] > 0.1){
    SceLC <- SceLC[Average_Counts > -1 & PropExpressingCells > 0.1, ]
    rownames(SceLC) <- genenames[Average_Counts > -1 & PropExpressingCells > 0.1]
  } else {
    SceLC <- rbind(SceLC[Average_Counts > -1 & PropExpressingCells > 0.1, ], SceLC["hSNCA-A53T",])
    rownames(SceLC) <- c(genenames[Average_Counts > -1 & PropExpressingCells > 0.1], "hSNCA-A53T")
  }
  rm(.Random.seed)
  set.seed(0)
  rm(SeuObj)
  colData(SceLC)$Treat <- factor(x = colData(SceLC)$Treat)
  colData(SceLC)$Age <- factor(x = colData(SceLC)$Age)
  colData(SceLC)$orig.ident <- factor(x = colData(SceLC)$orig.ident)
  blas_set_num_threads(1)
  assay(SceLC) <- as.matrix(assay(SceLC))
  SceLC <- scran::convertTo(SceLC, type = "DESeq2")
  weights <- computeObservationalWeights(zinbFit(SceLC, X="~Treat", commondispersion = TRUE, epsilon=1e12), counts(SceLC))
  assays(SceLC,  withDimnames = FALSE)[["weights"]] <- weights
  SceLC <- scran::computeSumFactors(SceLC)#, clusters = clusters) # Compute sizeFactors for each cell avoiding biases because of dropouts (using deconvolution method by scran).
  design(SceLC) <- formula(~Treat)
  SceLC <- DESeq2::estimateDispersions(object = SceLC, fitType='local') # fitType "local" because the algorithm couldn't find parameters a and b to fit a function between mean and variance (and that is because is preatty plain the relationship)
  SceLC <- nbinomLRT(SceLC, reduced= ~1, minmu = 1e-6)
```

#### log2FC of PD/Ctrl in jmo 3rd:

```{r}
eff_PD_jmo_3rd <- DESeq2::results(object = SceLC, contrast = list(c("Treat_PD_vs_Ctrl")), alpha = 0.05, independentFiltering=F, cooksCutoff=F ) %>% data.frame() %>% mutate(Gene = rownames(.))
PD <- coef(SceLC)[,c("Intercept", "Treat_PD_vs_Ctrl" )] %>%   rowSums()
  Ctrl <- coef(SceLC)[,c("Intercept")]
  log2FC <- log2(2^(PD)/2^(Ctrl))
  FC_df <- data.frame(log2FC, Gene = rownames(coef(SceLC)))
  FC_df_jmo_3rd4th <- inner_join(eff_PD_jmo_3rd, FC_df)
  GeneLab <- if_else(FC_df_jmo_3rd4th$Gene == "hSNCA-A53T", "hSNCA-A53T", "NA")
  GeneLab[GeneLab == "NA"] <- NA
  FC_df_jmo_3rd4th$GeneLab <- GeneLab
```


## Volcano plots

#### Threshold in padj-value = 0.05:

```{r}
    FC_df_jmo_3rd4th %>% FCfun(0.5, 0.05) + 
    labs(title = paste0("The effect of PD in the ", j), subtitle = "In the 3rd4th DS")
```

How many of these are considered replicate specific genes?
```{r}
# Import replicate specific genes from corresponding contrast
genes_replicate_specific <- readRDS(paste0("Outputs/", date, "_genes_replicate_specific_clust", m, "_", j, "_LC.rds"))

# Erase these genes replicate specific from the volcano plot.
FC_df_jmo_3rd4th %>% FCfun2(0.5, 0.05, genes_replicate_specific) + 
labs(title = paste0("The effect of PD in the ", j), subtitle = "No replicate specific")
```

How many of these are ALSO not differentially expressed in the full model of first analysis?
```{r}
# Erase these genes replicate specific from the volcano plot.
FC_df_jmo_3rd4th %>% FCfun2(0.5, 0.05, unique(c(genes_replicate_specific, nonDE_genes))) + 
labs(title = paste0("The effect of PD in the ", j), subtitle = "No replicate specific neither nonDE on full model")
```

If I use a lower log2FC threshold:

```{r}
    FC_df_jmo_3rd4th  %>% FCfun(0.3, 0.05) + 
    labs(title = paste0("The effect of PD in the ", j), subtitle = "In the 3rd4th DS")
```

Without replicate specific genes:
```{r}
# Erase these genes replicate specific from the volcano plot.
FC_df_jmo_3rd4th %>% FCfun2(0.3, 0.05, genes_replicate_specific) + 
labs(title = paste0("The effect of PD in the ", j), subtitle = "No replicate specific")
```

How many of these are ALSO not differentially expressed in the full model of first analysis?
```{r}
# Erase these genes replicate specific from the volcano plot.
FC_df_jmo_3rd4th %>% FCfun2(0.3, 0.05, unique(c(genes_replicate_specific, nonDE_genes))) + 
labs(title = paste0("The effect of PD in the ", j), subtitle = "No replicate specific neither nonDE on full model")
```

```{r}
# Saving these table
write_csv(filter(FC_df_jmo_3rd4th), paste0("Outputs/", date, "_FCmat_posthoc_nofilter_log2FC_clust", m, "_", j, "_LC.csv"))
write_csv(filter(FC_df_jmo_3rd4th, !(Gene %in% nonDE_genes)), paste0("Outputs/", date, "_FCmat_posthoc_nonDEfilter_log2FC_clust", m, "_", j, "_LC.csv"))
write_csv(filter(FC_df_jmo_3rd4th, !(Gene %in% genes_replicate_specific)), paste0("Outputs/", date, "_FCmat_replicatefilter_official_log2FC_clust", m, "_", j, "_LC.csv"))
write_csv(filter(FC_df_jmo_3rd4th, !(Gene %in% c(genes_replicate_specific, nonDE_genes))), paste0("Outputs/", date, "_FCmat_nonDEreplicatefilter_log2FC_clust", m, "_", j, "_LC.csv"))
```

Different types of not considered genes

```{r}
DESeq2_output_sig <- filter(DESeq2_output, padj <= 0.05)
comparison <- inner_join(FC_df_jmo_3rd4th, DESeq2_output_sig, by = "Gene")
filter(comparison, padj.x > 0.05, Gene %in% genes_replicate_specific)$Gene # 7, entre ellos hSNCA-A53T
comparison2 <- filter(comparison, !(Gene %in% genes_replicate_specific))
filter(comparison2, padj.x <= 0.05)$Gene #39
filter(comparison2, padj.x > 0.05)$Gene #12 
filter(comparison2, abs(log2FC3rd) < 0.3, abs(log2FC) > 0.3)$Gene # 9 
filter(comparison2, abs(log2FC4th) < 0.3, abs(log2FC) > 0.3)$Gene # 7 distintos
filter(comparison2, log2FC3rd < -0.3, log2FC > 0.3)$Gene # 0 
filter(comparison2, log2FC3rd > 0.3, log2FC < -0.3)$Gene # 3
filter(comparison2, log2FC4th < -0.3, log2FC > 0.3)$Gene # 0 
filter(comparison2, log2FC4th > 0.3, log2FC < -0.3)$Gene # 1
```

