---
title: "Data Transfer NE-LC"
author: "Camila Arcuschin; Ignacio Schor"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    theme: cerulean
    code_folding: hide
out.width: '30%'
out.heigh: '100%'
fig.width: '100%'
fig.height: '100%'
dpi: 100
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = TRUE, warning = FALSE, cache=F, dev = c('png', 'pdf'))
knitr::opts_knit$set(root.dir = "/data_husihuilke/shared/SMB/carcu/Parkinsons_project/")
```

```{r, include=FALSE}
library(Seurat)
library(SeuratDisk)
library(dplyr)
library(magrittr)
library(ggplot2)
library(cowplot)
library(future)
#register(MulticoreParam(10))
plan(multisession, workers = 6)
# plan("multiprocess", workers = 6)
options(future.globals.maxSize = 4000 * 1024^2)
options(ggrepel.max.overlaps = Inf)
```

plotear tmb los clusters en la transferencia al UMAP de query, probar tambien hacer un rpca para el transfer label y  plotear tambien los grupos de pagoda cuando lo hago sin glia

```{r}
CellType_colors <- c("GABA/Glut.ergic" = "#d11141","GABAergic" = "#f37735",  "Glutamatergic_Slc17a6" = "#00b159", "Glutamatergic_Slc17a7" = "#AACC00", "NEergic" = "#00aedb", "Slc17a6/Barrington?" = "#B56576")
colors_n <- readRDS("Outputs/231129_QC_first_part_tidy_cluster_colors_LC.rds")
#colors_n <- colors_n[as.character(c(0:14))[-13]]

my_theme <- theme_void(base_size = 6) + theme(title = element_text(face = "bold"))
```


## RNA(after-integration)-RNA-LogNormalize-PCAproject (reference projects into the query)

Import integrated data and reference data

```{r}
SeuInt <- readRDS("Outputs/231129_SeuInt_3rd4th_fil_int_mayt1000lowt60000_lowtInfrb_24clusters.rds")

SeuInt <- subset(SeuInt, subset = Treat == "Ctrl")
#SeuInt <- subset(SeuInt, subset = Treat == "Ctrl", idents = c(0,1,6,7,9,10,11,13,14))

SeuRef <-readRDS("/data_husihuilke/shared/SMB/carcu/Caramia_2023/data/231018_SeuratProject_integrated_allLC.rds")
DimPlot(SeuRef, cols = CellType_colors)
DefaultAssay(SeuRef) <- "RNA"
SeuRef <- NormalizeData(SeuRef,
                normalization.method = "LogNormalize",
                scale.factor = 10000) %>%
  FindVariableFeatures(
    selection.method = "vst",
    nfeatures = 2000,
    loess.span = 0.6
  )
SeuRef <- ScaleData(SeuRef) %>%
  RunPCA(npcs = 30, reduction.name = "pca.RNA")

```

After finding anchors, we use the TransferData() function to classify the query cells based on reference data (a vector of reference cell type labels). TransferData() returns a matrix with predicted IDs and prediction scores, which we can add to the query metadata.

```{r}
anchors <- FindTransferAnchors(
  reference = SeuRef,
  reference.assay = "RNA",
  query = SeuInt,
  query.assay = "RNA",
  dims = 1:30,
  reference.reduction = "pca.RNA",
  normalization.method = "LogNormalize"
)
predictions <- TransferData(anchorset = anchors,
                            refdata = SeuRef$CellType,
                            dims = 1:30)
SeuInt <- AddMetaData(SeuInt, metadata = predictions)
```

```{r}
DefaultAssay(SeuInt) <- "INTE"
a <- DimPlot(SeuInt, label = T, cols = colors_n) + theme(legend.position = "none")
b <- DimPlot(SeuInt, group.by = "predicted.id", label = T, cols = CellType_colors, order = F) + theme(legend.position = "none")
a + b

library(ggrepel)

DefaultAssay(SeuRef) <- "integrated"
SeuRef <-  RunUMAP(
    SeuRef,
    dims = 1:30,
    reduction = "pca",
    return.model = TRUE
  )
SeuIntq <-  MapQuery(
    anchorset = anchors,
    reference = SeuRef,
    query = SeuInt,
    refdata = list(celltype = "CellType"),
    reference.reduction = "pca.RNA", #"pca.RNA"
    reduction.model = "umap"
  )
p1 <-  DimPlot(
    SeuRef,
    reduction = "umap",
    group.by = "CellType",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = CellType_colors,
    pt.size = 0.3,
    order = F
  ) + my_theme + theme(legend.position = "none") + ggtitle("Reference annotations")
p2 <-  DimPlot(
    SeuIntq,
    reduction = "ref.umap",
    group.by = "predicted.id",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = CellType_colors,
    pt.size = 0.3,
    order = F
  ) + my_theme + theme(legend.position = "none") + ggtitle("Query transferred labels") 
p3 <-  DimPlot(
    SeuIntq,
    reduction = "ref.umap",
    group.by = "seurat_clusters",
    label = T,
    label.size = 3,
    repel = TRUE,
    cols = colors_n,
    pt.size = 0.3,
    order = F,
    shuffle = T
  ) + my_theme + ggtitle("Clusters") + theme(legend.position = "none")
p6 <-  DimPlot(
    SeuRef,
    reduction = "umap",
    group.by = "seurat_clusters",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    pt.size = 0.3,
    order = F,
    shuffle = T
  ) + my_theme + theme(legend.position = "none") + ggtitle("Reference annotations")
plot_grid(p1, p2, p3, p6, ncol = 3, nrow = 2)
```

## RNA(but pca after-integration)-RNA-LogNormalize-RPCA (one projects into the other and viceversa for the query RNA, that is why I choose it)


```{r}
anchors <- FindTransferAnchors(
  reference = SeuRef,
  reference.assay = "RNA",
  reduction = "rpca",
  reference.reduction = "pca", #Even in reciprocal PCA: # pcaproject is used as the weight.matrix in MapQuery
  query = SeuInt,
  query.assay = "RNA",
  dims = 1:30,
  normalization.method = "LogNormalize"
)
predictions <- TransferData(anchorset = anchors,
                            refdata = SeuRef$CellType,
                            dims = 1:30)
saveRDS(predictions, "Outputs/240222_DataTransfer_NELC_predictions_ctrl.rds")
SeuInt <- AddMetaData(SeuInt, metadata = predictions)
```

```{r}
DefaultAssay(SeuInt) <- "INTE"
a <- DimPlot(SeuInt, label = T, cols = colors_n) + theme(legend.position = "none")
b <- DimPlot(SeuInt, group.by = "predicted.id", label = F, cols = CellType_colors, order = F) + theme(legend.position = "none")
a + b

DefaultAssay(SeuRef) <- "integrated"
SeuRef <-  RunUMAP(
    SeuRef,
    dims = 1:30,
    reduction = "pca",
    return.model = TRUE
  )
SeuIntq <-  MapQuery(
    anchorset = anchors,
    reference = SeuRef,
    query = SeuInt,
    refdata = list(celltype = "CellType"),
    # By default: slot(object = anchors, name = "command")$reference #rpca.ref.l2. #Even in reciprocal PCA: # pcaproject is used as the weight.matrix in MapQuery.
    # So it only use the anchors of the RPCA but using the projection of the query on the reference to plot the query. 
    reduction.model = "umap"
  )
# all(SeuIntq$predicted.celltype.score == SeuInt$prediction.score.max) # T

p1 <-  DimPlot(
    SeuRef,
    reduction = "umap",
    group.by = "CellType",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = CellType_colors,
    pt.size = 0.3,
    order = F
  ) + my_theme + theme(legend.position = "none") + ggtitle("Reference annotations")
p2 <-  DimPlot(
    SeuIntq,
    reduction = "ref.umap",
    group.by = "predicted.id",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = CellType_colors,
    pt.size = 0.3,
    order = F
  ) + my_theme + theme(legend.position = "none") + ggtitle("Query transferred labels") 
p3 <-  DimPlot(
    SeuIntq,
    reduction = "ref.umap",
    group.by = "seurat_clusters",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = colors_n,
    pt.size = 0.3,
    order = F
  ) + my_theme + theme(legend.position = "none") + ggtitle("Clusters") 
p6 <-  DimPlot(
    SeuRef,
    reduction = "umap",
    group.by = "seurat_clusters",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    pt.size = 0.3,
    order = F
  ) + my_theme + theme(legend.position = "none") + ggtitle("Reference annotations")
plot_grid(p1, p2, p3, p6, ncol = 3, nrow = 2)

saveRDS(SeuInt, "Outputs/231024_SeuInt_clean_clusters_neurons_cell_annotation.rds")
```


## integrated-INTE.SCT(first-normalized-then-integrated)-SCT normalization-Recompute residuals

```{r}
SeuInt_o <- readRDS("Outputs/231129_SeuInt_3rd4th_fil_int_mayt1000lowt60000_lowtInfrb_24clusters.rds")
SeuInt_o <- subset(SeuInt_o, subset = Treat == "Ctrl")
#SeuInt_o <- subset(SeuInt_o, subset = Treat == "Ctrl", idents = c(0,1,6,7,9,10,11,13,14))

SeuInt <- SeuInt_o
SeuInt_list <- SplitObject(SeuInt, split.by = "orig.ident")
for (i in 1:length(SeuInt_list)) {
  SeuInt_list[[i]] <-
    SCTransform(
      SeuInt_list[[i]],
      assay = "RNA",
      return.only.var.genes = FALSE,
      verbose = FALSE,
      variable.features.n = 10000
    )
  SeuInt_list[[i]] <-
    RunPCA(SeuInt_list[[i]], npcs = 30, reduction.name = "pca.SCT")
}
Int.features <-
  SelectIntegrationFeatures(object.list = SeuInt_list, nfeatures = 10000)
SeuInt_list <-
  PrepSCTIntegration(object.list = SeuInt_list, anchor.features = Int.features)
Int.anchors <-
  FindIntegrationAnchors(
    object.list = SeuInt_list,
    anchor.features = Int.features,
    normalization.method = "SCT"
  )
SeuInt <-
  IntegrateData(
    anchorset = Int.anchors,
    normalization.method = "SCT",
    new.assay.name = "INTE.SCT"
  )
DefaultAssay(SeuInt) <- "INTE.SCT"
SeuInt <- RunPCA(SeuInt, npcs = 30, verbose = FALSE)

SeuRef <-
  readRDS("/data_husihuilke/shared/SMB/carcu/Caramia_2023/data/231018_SeuratProject_integrated_allLC.rds")
DefaultAssay(SeuRef) <- "integrated"
rm(SeuInt_list)
gc()
```

```{r}
anchors <- FindTransferAnchors(
    reference = SeuRef,
    reference.assay = "integrated",
    query = SeuInt,
    query.assay = "INTE.SCT",
    dims = 1:30,
    normalization.method = "SCT",
    reference.reduction = "pca"
  )
predictions <- TransferData(anchorset = anchors,
               refdata = SeuRef$CellType,
               dims = 1:30)
SeuInt <- AddMetaData(SeuInt, metadata = predictions)
```

```{r}
DefaultAssay(SeuInt) <- "INTE"
SeuInt@reductions$umap <- SeuInt_o@reductions$umap
a <- DimPlot(SeuInt, label = T, cols = colors_n, order = F) + theme(legend.position = "none")
b <- DimPlot(SeuInt, group.by = "predicted.id", label = T, cols = CellType_colors, order = F) + theme(legend.position = "none")
a + b

DefaultAssay(SeuRef) <- "integrated"
SeuRef <-  RunUMAP(
    SeuRef,
    dims = 1:30,
    reduction = "pca",
    return.model = TRUE
  )
DimPlot(SeuRef, cols = CellType_colors)
SeuIntq <-  MapQuery(
    anchorset = anchors,
    reference = SeuRef,
    query = SeuInt,
    refdata = list(celltype = "CellType"),
    reference.reduction = "pca",
    reduction.model = "umap"
  )

p1 <-  DimPlot(
    SeuRef,
    reduction = "umap",
    group.by = "CellType",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = CellType_colors,
    pt.size = 0.3,
    order = F
  ) + my_theme + theme(legend.position = "none") + ggtitle("Reference annotations")
p2 <-  DimPlot(
    SeuIntq,
    reduction = "ref.umap",
    group.by = "predicted.id",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = CellType_colors,
    pt.size = 0.3,
    order = F
   ) + my_theme + theme(legend.position = "none") + ggtitle("Query transferred labels")
p3 <-  DimPlot(
    SeuIntq,
    reduction = "ref.umap",
    group.by = "seurat_clusters",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = colors_n,
    pt.size = 0.3,
    order = F
   ) + my_theme + theme(legend.position = "none") + ggtitle("Clusters")
p6 <-  DimPlot(
    SeuRef,
    reduction = "umap",
    group.by = "seurat_clusters",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    pt.size = 0.3,
    order = F
  ) + my_theme + theme(legend.position = "none") + ggtitle("Reference annotations")
plot_grid(p1, p2, p3, p6, ncol = 3, nrow = 2)
```


## integrated-INTE.SCT(first-normalized-then-integrated)-SCT Normalization- without Recompute residuals



```{r}
anchors <- FindTransferAnchors(
    reference = SeuRef,
    reference.assay = "integrated",
    query = SeuInt,
    query.assay = "INTE.SCT",
    recompute.residuals = F,
    dims = 1:30,
    normalization.method = "SCT",
    reference.reduction = "pca"
  )
predictions <- TransferData(anchorset = anchors,
               refdata = SeuRef$CellType,
               dims = 1:30)
SeuInt <- AddMetaData(SeuInt, metadata = predictions)
```

```{r}
DefaultAssay(SeuInt) <- "INTE"
SeuInt@reductions$umap <- SeuInt_o@reductions$umap
a <- DimPlot(SeuInt, label = T, cols = colors_n, order = F) + theme(legend.position = "none")
b <- DimPlot(SeuInt, group.by = "predicted.id", label = T, cols = CellType_colors, order = F) + theme(legend.position = "none")
a + b

DefaultAssay(SeuRef) <- "integrated"
SeuRef <-  RunUMAP(
    SeuRef,
    dims = 1:30,
    reduction = "pca",
    return.model = TRUE
  )
DimPlot(SeuRef, cols = CellType_colors)
SeuIntq <-  MapQuery(
    anchorset = anchors,
    reference = SeuRef,
    query = SeuInt,
    refdata = list(celltype = "CellType"),
    reference.reduction = "pca",
    reduction.model = "umap"
  )

p1 <-  DimPlot(
    SeuRef,
    reduction = "umap",
    group.by = "CellType",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = CellType_colors,
    pt.size = 0.3,
    order = F
  ) + my_theme + theme(legend.position = "none") + ggtitle("Reference annotations")
p2 <-  DimPlot(
    SeuIntq,
    reduction = "ref.umap",
    group.by = "predicted.id",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = CellType_colors,
    pt.size = 0.3,
    order = F
   ) + my_theme + theme(legend.position = "none") + ggtitle("Query transferred labels")
p3 <-  DimPlot(
    SeuIntq,
    reduction = "ref.umap",
    group.by = "seurat_clusters",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = colors_n,
    pt.size = 0.3,
    order = F
   ) + my_theme + theme(legend.position = "none") + ggtitle("Clusters")
p6 <-  DimPlot(
    SeuRef,
    reduction = "umap",
    group.by = "seurat_clusters",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    pt.size = 0.3,
    order = F
  ) + my_theme + theme(legend.position = "none") + ggtitle("Reference annotations")
plot_grid(p1, p2, p3, p6, ncol = 3, nrow = 2)
```

## INTE(LogNormalize-then-integrate)-INTE-LogNormalize

```{r}
SeuInt <- readRDS("Outputs/231129_SeuInt_3rd4th_fil_int_mayt1000lowt60000_lowtInfrb_24clusters.rds")
SeuInt <- subset(SeuInt, subset = Treat == "Ctrl")
#SeuInt <- subset(SeuInt, subset = Treat == "Ctrl", idents = c(0,1,6,7,9,10,11,13,14))
DefaultAssay(SeuInt) <- "INTE"

SeuRef <- readRDS("/data_husihuilke/shared/SMB/carcu/Caramia_2023/data/231018_SeuratProject_integrated_allLC.rds")
DefaultAssay(SeuRef) <- "RNA"
SeuRef_list <- SplitObject(SeuRef, split.by = "id")
for (i in 1:length(SeuRef_list)) {
  DefaultAssay(SeuRef_list[[i]]) <- "RNA"
  SeuRef_list[[i]] <-  NormalizeData(SeuRef_list[[i]],
                  normalization.method = "LogNormalize",
                  scale.factor = 10000) %>%
    FindVariableFeatures(
      selection.method = "vst",
      nfeatures = 2000,
      loess.span = 0.6
    )
  SeuRef_list[[i]] <- ScaleData(SeuRef_list[[i]]) %>%
    RunPCA(npcs = 30)
}
SeuRefFeatures <- SelectIntegrationFeatures(SeuRef_list)
Ref.anchors <- FindIntegrationAnchors(
    SeuRef_list,
    reduction = "cca",
    dims = 1:30,
    anchor.features = SeuRefFeatures,
    scale = F,
    normalization.method = c("LogNormalize")
  )
# LCmark <-
#   c("Dbh",
#     "Th",
#     "Slc6a2",
#     "Chrna7",
#     "Maoa",
#     "Rlim",
#     "Sec23a",
#     "Ppp2r2d",
#     "Usp34")
# Neuron <- c("Rbfox3") # General neurons
# Ex <- c("Slc17a6", "Slc17a7") # Excitatory neurons
# Inh <- c("Gad1") # Inhibitory neuron
SeuRef <-  IntegrateData(
  anchorset = Ref.anchors,
  dims = 1:30,
  new.assay.name = "INTE",
  features.to.integrate = c(
    SeuRefFeatures
    # LCmark,
    # Neuron,
    # Ex,
    # Inh
  )
)
SeuRef <- ScaleData(SeuRef, npcs = 30, verbose = FALSE)
SeuRef <- RunPCA(SeuRef, npcs = 30, verbose = FALSE)
SeuRef <-
  RunUMAP(
    SeuRef,
    reduction = "pca",
    dims = 1:30,
    return.model = TRUE
  )
DefaultAssay(SeuRef) <- "INTE"
rm(SeuRef_list)
gc()
```

```{r}
anchors <-  FindTransferAnchors(
    reference = SeuRef,
    reference.assay = "INTE",
    query = SeuInt,
    query.assay = "INTE",
    dims = 1:30,
    normalization.method = "LogNormalize",
    reference.reduction = "pca",
    recompute.residuals = F
  )
predictions <- TransferData(anchorset = anchors,
               refdata = SeuRef$CellType,
               dims = 1:30)
SeuInt <- AddMetaData(SeuInt, metadata = predictions)
```

```{r}
DefaultAssay(SeuInt) <- "INTE"
a <- DimPlot(SeuInt, label = T, cols = colors_n, order = F) + theme(legend.position = "none")
b <- DimPlot(SeuInt, group.by = "predicted.id", label = T, cols = CellType_colors, order = F) + theme(legend.position = "none")
a + b

DefaultAssay(SeuRef) <- "INTE"
#SeuRef <- RunUMAP(SeuRef, dims = 1:30, reduction = "pca", return.model = TRUE)
DimPlot(SeuRef, cols = CellType_colors)
SeuIntq <-  MapQuery(
    anchorset = anchors,
    reference = SeuRef,
    query = SeuInt,
    refdata = list(celltype = "CellType"),
    reference.reduction = "pca",
    reduction.model = "umap"
  )

p1 <-  DimPlot(
    SeuRef,
    reduction = "umap",
    group.by = "CellType",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = CellType_colors,
    pt.size = 0.3,
    order = F
   ) + my_theme + theme(legend.position = "none") + ggtitle("Reference annotations")
p2 <-  DimPlot(
    SeuIntq,
    reduction = "ref.umap",
    group.by = "predicted.id",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = CellType_colors,
    pt.size = 0.3,
    order = F
   ) + my_theme + theme(legend.position = "none") + ggtitle("Query transferred labels")
p3 <-  DimPlot(
    SeuIntq,
    reduction = "ref.umap",
    group.by = "seurat_clusters",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = colors_n,
    pt.size = 0.3,
    order = F
   ) + my_theme + theme(legend.position = "none") + ggtitle("Clusters")
p6 <-  DimPlot(
    SeuRef,
    reduction = "umap",
    group.by = "seurat_clusters",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    pt.size = 0.3,
    order = F
   ) + my_theme + my_theme + theme(legend.position = "none") + ggtitle("Reference annotations")
plot_grid(p1, p2, p3, p6, ncol = 3, nrow = 2)
```

## INTE-INTE-LogNormalize-RenormalizingBothBeforeIntegrate

```{r}
SeuInt_o <- readRDS("Outputs/231129_SeuInt_3rd4th_fil_int_mayt1000lowt60000_lowtInfrb_24clusters.rds")
SeuInt_o <- subset(SeuInt_o, subset = Treat == "Ctrl")
#SeuInt_o <- subset(SeuInt_o, subset = Treat == "Ctrl", idents = c(0,1,6,7,9,10,11,13,14))

SeuInt <- SeuInt_o
DefaultAssay(SeuInt) <- "RNA"
SeuInt_list <- SplitObject(SeuInt, split.by = "orig.ident")
for (i in 1:length(SeuInt_list)) {
  DefaultAssay(SeuInt_list[[i]]) <- "RNA"
  SeuInt_list[[i]] <-  NormalizeData(SeuInt_list[[i]],
                  normalization.method = "LogNormalize",
                  scale.factor = 10000) %>%
    FindVariableFeatures(
      selection.method = "vst",
      nfeatures = 2000,
      loess.span = 0.6
    )
  SeuInt_list[[i]] <- ScaleData(SeuInt_list[[i]]) %>%
    RunPCA(npcs = 30)
}
obj_dims = 30
SeuIntFeatures <- SelectIntegrationFeatures(SeuInt_list) 
anchors <- FindIntegrationAnchors(SeuInt_list, reduction = "rpca", reference = 1,  dims = 1:obj_dims, anchor.features = SeuIntFeatures, scale = F, k.anchor = 10, k.filter = 100,  k.score = 20, max.features = 100) 
LCmark <- c("Dbh", "Th", "Slc6a2","Chrna7", "Maoa", "Rlim", "Sec23a", "Ppp2r2d", "Usp34") 
Neuron <- c("Rbfox3") 
Ex <- c("Slc17a6", "Slc17a7") 
Inh <- c("Gad1") 
Astrocytes <- c("Gfap", "Gja1") 
Microglia <- c("C1qb") 
OPC <- c("Olig1", "Tnr")
Oligodendrocyte <- c("Tfrc", "Plp1", "Mbp") 
Fibro <- c("Dcn") 
Capil <- c("Flt1", "Rgs5")
markers <- c("EGFP-KASH", "hSNCA-A53T", "mCherry",  "Snca")
SeuInt <- IntegrateData(anchorset = anchors, dims = 1:obj_dims, k.weight = 20, new.assay.name = "INTE", features.to.integrate = c(SeuIntFeatures, LCmark, Neuron, Ex, Inh, Fibro, Astrocytes, Microglia, Oligodendrocyte, OPC, Capil, markers))
DefaultAssay(SeuInt) <- "INTE"
rm(SeuInt_list)
gc()


SeuRef <- readRDS("/data_husihuilke/shared/SMB/carcu/Caramia_2023/data/231018_SeuratProject_integrated_allLC.rds")
DefaultAssay(SeuRef) <- "RNA"
SeuRef_list <- SplitObject(SeuRef, split.by = "id")
for (i in 1:length(SeuRef_list)) {
  DefaultAssay(SeuRef_list[[i]]) <- "RNA"
  SeuRef_list[[i]] <-  NormalizeData(SeuRef_list[[i]],
                  normalization.method = "LogNormalize",
                  scale.factor = 10000) %>%
    FindVariableFeatures(
      selection.method = "vst",
      nfeatures = 2000,
      loess.span = 0.6
    )
  SeuRef_list[[i]] <- ScaleData(SeuRef_list[[i]]) %>%
    RunPCA(npcs = 30)
}
SeuRefFeatures <- SelectIntegrationFeatures(SeuRef_list)
Ref.anchors <- FindIntegrationAnchors(
    SeuRef_list,
    reduction = "cca",
    dims = 1:30,
    anchor.features = SeuRefFeatures,
    scale = F,
    normalization.method = c("LogNormalize")
  )
SeuRef <-  IntegrateData(
  anchorset = Ref.anchors,
  dims = 1:30,
  new.assay.name = "INTE",
  features.to.integrate = c(
    SeuRefFeatures
    # LCmark,
    # Neuron,
    # Ex,
    # Inh
  )
)
SeuRef <- ScaleData(SeuRef, npcs = 30, verbose = FALSE)
SeuRef <- RunPCA(SeuRef, npcs = 30, verbose = FALSE)
SeuRef <-
  RunUMAP(
    SeuRef,
    reduction = "pca",
    dims = 1:30,
    return.model = TRUE
  )
DefaultAssay(SeuRef) <- "INTE"
rm(SeuRef_list)
gc()
```

```{r}
anchors <-  FindTransferAnchors(
    reference = SeuRef,
    reference.assay = "INTE",
    query = SeuInt,
    query.assay = "INTE",
    dims = 1:30,
    normalization.method = "LogNormalize",
    reference.reduction = "pca",
    recompute.residuals = F
  )
predictions <- TransferData(anchorset = anchors,
               refdata = SeuRef$CellType,
               dims = 1:30)
SeuInt <- AddMetaData(SeuInt, metadata = predictions)
```

```{r}
DefaultAssay(SeuInt) <- "INTE"
SeuInt@reductions$umap <- SeuInt_o@reductions$umap
a <- DimPlot(SeuInt, label = T, cols = colors_n, order = F) + theme(legend.position = "none")
b <- DimPlot(SeuInt, group.by = "predicted.id", label = T, cols = CellType_colors, order = F) + theme(legend.position = "none")
a + b

DefaultAssay(SeuRef) <- "INTE"
#SeuRef <- RunUMAP(SeuRef, dims = 1:30, reduction = "pca", return.model = TRUE)
DimPlot(SeuRef, cols = CellType_colors)
SeuIntq <-  MapQuery(
    anchorset = anchors,
    reference = SeuRef,
    query = SeuInt,
    refdata = list(celltype = "CellType"),
    reference.reduction = "pca",
    reduction.model = "umap"
  )

p1 <-  DimPlot(
    SeuRef,
    reduction = "umap",
    group.by = "CellType",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = CellType_colors,
    pt.size = 0.3,
    order = F
   ) + my_theme + theme(legend.position = "none") + ggtitle("Reference annotations")
p2 <-  DimPlot(
    SeuIntq,
    reduction = "ref.umap",
    group.by = "predicted.id",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = CellType_colors,
    pt.size = 0.3,
    order = F
   ) + my_theme + theme(legend.position = "none") + ggtitle("Query transferred labels")
p3 <-  DimPlot(
    SeuIntq,
    reduction = "ref.umap",
    group.by = "seurat_clusters",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = colors_n,
    pt.size = 0.3,
    order = F
   ) + my_theme + theme(legend.position = "none") + ggtitle("Clusters")
p6 <-  DimPlot(
    SeuRef,
    reduction = "umap",
    group.by = "seurat_clusters",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    pt.size = 0.3,
    order = F
  ) + my_theme + theme(legend.position = "none") + ggtitle("Reference annotations")
plot_grid(p1, p2, p3, p6, ncol = 3, nrow = 2)
```

## INTE-RNA-SCT-RecomputeResiduals

Equivalent to run 'SCTransform()' on the query data (integrate-then-SCTransform) but using the reference data for the 'reference.SCT.model'parameter. Also 'recompute.residuals' is always true by default when normalization method is "SCT".

```{r}
SeuInt <- readRDS("Outputs/231129_SeuInt_3rd4th_fil_int_mayt1000lowt60000_lowtInfrb_24clusters.rds")
SeuInt <- subset(SeuInt, subset = Treat == "Ctrl")
#SeuInt <- subset(SeuInt_o, subset = Treat == "Ctrl", idents = c(0,1,6,7,9,10,11,13,14))
DefaultAssay(SeuInt) <- "RNA"

SeuInt <- FindVariableFeatures(SeuInt, nfeatures = 10000)

SeuRef <- readRDS("/data_husihuilke/shared/SMB/carcu/Caramia_2023/data/231018_SeuratProject_integrated_allLC.rds")
DefaultAssay(SeuRef) <- "integrated"
```

```{r}
anchors <-  FindTransferAnchors(
    reference = SeuRef,
    reference.assay = "integrated",
    query = SeuInt,
    query.assay = "RNA",
    dims = 1:30,
    features = VariableFeatures(SeuInt)[VariableFeatures(SeuInt) %in% VariableFeatures(SeuRef)],
    normalization.method = "SCT",
    reference.reduction = "pca",
    recompute.residuals = T
  )
predictions <- TransferData(anchorset = anchors,
               refdata = SeuRef$CellType,
               dims = 1:30)
SeuInt <- AddMetaData(SeuInt, metadata = predictions)
```

```{r}
DefaultAssay(SeuInt) <- "INTE"
a <- DimPlot(SeuInt, label = T, cols = colors_n, order = F) + theme(legend.position = "none")
b <- DimPlot(SeuInt, group.by = "predicted.id", label = T, cols = CellType_colors, order = F) + theme(legend.position = "none")
a + b

DefaultAssay(SeuRef) <- "integrated"
SeuRef <- RunUMAP(SeuRef, dims = 1:30, reduction = "pca", return.model = TRUE)
DimPlot(SeuRef, cols = CellType_colors)
SeuIntq <-  MapQuery(
    anchorset = anchors,
    reference = SeuRef,
    query = SeuInt,
    refdata = list(celltype = "CellType"),
    reference.reduction = "pca",
    reduction.model = "umap"
  )

p1 <-  DimPlot(
    SeuRef,
    reduction = "umap",
    group.by = "CellType",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = CellType_colors,
    pt.size = 0.3,
    order = F
   ) + my_theme + theme(legend.position = "none") + ggtitle("Reference annotations")
p2 <-  DimPlot(
    SeuIntq,
    reduction = "ref.umap",
    group.by = "predicted.id",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = CellType_colors,
    pt.size = 0.3,
    order = F
   ) + my_theme + theme(legend.position = "none") + ggtitle("Query transferred labels")
p3 <-  DimPlot(
    SeuIntq,
    reduction = "ref.umap",
    group.by = "seurat_clusters",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    cols = colors_n,
    pt.size = 0.3,
    order = F
   ) + my_theme + theme(legend.position = "none") + ggtitle("Clusters")
p6 <-  DimPlot(
    SeuRef,
    reduction = "umap",
    group.by = "seurat_clusters",
    label = TRUE,
    label.size = 3,
    repel = TRUE,
    pt.size = 0.3,
    order = F
  ) + my_theme + theme(legend.position = "none") + ggtitle("Reference annotations")
plot_grid(p1, p2, p3, p6, ncol = 3, nrow = 2)
```
