---
title: "GSVA"
author: "Camila Arcuschin; Ignacio Schor"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    bibliography: references.bib
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = FALSE, warning = FALSE, dpi=500)
knitr::opts_knit$set(root.dir = "/data_husihuilke/shared/SMB/carcu/Parkinsons_project/")
```

```{r}
library(tidyverse)
library(GSVA)
library(gplots)
library(ggcorrplot)
library(GSA)
library(scales)
library(cowplot)
library(ggplot2)
library(ggplotify)
```

```{r Functions}
GeneSetCor <- function(cor.matrix){
  ggcorrplot(cor.matrix, hc.order = F, 
           type = "full", 
           lab = TRUE, 
           lab_size = 3, title="Correlation of Expression", 
           ggtheme=theme_bw,
           show.diag = T)+scale_fill_gradient2(low="darkblue", high="red", mid = "white", midpoint=0.5)}

# Legend of the heatmap
get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

# Legend of the heatmap
get_plot<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  plot <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[plot]]
  return(legend)
}

# Data preparation
PrepareData <- function(enrichMat, clust, ctype, ctype_u, cols){
  colData <- data.frame(clusters= clust, CellType= ctype)
  colors <- data.frame(CellType = ctype_u , Cell_colors = cols)
  colData <- left_join(colData, colors)
  colData <- colData[match(colnames(enrichMat), colData$clusters),]
}

PrepareDataH <- function(enrichMat, h_result, h_ord, ctype_u, cols){
  colData <- data.frame(clusters= names(h_result), KType= h_result)
  colors <- data.frame(KType= h_ord, CellType = ctype_u , Cell_colors = cols)
  colData <- left_join(colData, colors)
  colData <- colData[match(colnames(enrichMat), colData$clusters),]
  colData <- colData[-2] # delate the KType column
}

# Get legend
plot4legend <- function(df){
  ggplot(df) +
    geom_bar(mapping = aes(x = clusters, y = 1, fill = CellType), 
             stat = "identity", 
             width = 1.1) +
    scale_fill_manual(labels = unique(df[2:3])$CellType, values = unique(df[2:3])$Cell_colors) +
    guides(fill=guide_legend(ncol=2))
}

# Plot HM
PlotHM <- function(enrichMat, sideCols){
  heatmap.2 (t(enrichMat), #enrich_hm <- 
                          # dendrogram control
                          Rowv = TRUE,
                          Colv= TRUE,
                          col = "bluered",
                          breaks = unique(c(seq(-1, -0.3, length=51), seq(-0.3, 0.3, length=101), seq(0.3, 1, length=51))),
                          distfun = dist,
                          trace = "none",
                          hclustfun = hclust,
                          dendrogram = "row",
                          symm=F,symkey=F,symbreaks=T, scale="none",
                          lmat=rbind(c(5,4,0), c(3,2,1)), 
                            lhei=c(2,5),
                          lwid=c(1,4,0.20),
                          RowSideColors = sideCols$Cell_colors,
                          cexCol = 0.7,
                          cexRow = 1.2,
                          srtCol = 45)
  legend(y=1.2, x=.8, xpd=TRUE,     
    legend = unique(sideCols$CellType),
    col = unique(sideCols$Cell_colors), 
    lty= 1,             
    lwd = 5,           
    cex=.7
    )
}

```

Read the objects and Get the correlations between the different genesets inside each object

```{r}
# from file: "GeneSet_generator_usingPublicData.R"
date = 231007
```

CellMarker cells (Mouse)
```{r}
GeneSet_CellMarkerMouse <- readRDS(paste0("Outputs/", date, "_GeneSet_CellMarkerMouse.rds"))
Ollaps_general <- computeGeneSetsOverlap(GeneSet_CellMarkerMouse, unique(unlist(GeneSet_CellMarkerMouse)))
GeneSetCor(Ollaps_general)
ggsave(filename = paste0("Outputs/", date, "_GeneSet_CellMarkerMouse_cor.png"))
```

CellMarker cells (Human)

```{r}
GeneSet_CellMarkerHuman <- readRDS(paste0("Outputs/", date, "_GeneSet_CellMarkerHuman.rds"))
Ollaps_general <- computeGeneSetsOverlap(GeneSet_CellMarkerHuman, unique(unlist(GeneSet_CellMarkerHuman)))
GeneSetCor(Ollaps_general)
ggsave(filename = paste0("Outputs/", date, "_GeneSet_CellMarkerHuman_cor.png"))
```

Panglao (Mouse)
```{r}
GeneSet_PanglaoMouse <- readRDS(paste0("Outputs/", date, "_GeneSet_PanglaoMouse.rds"))
Ollaps_Allen  <- computeGeneSetsOverlap(GeneSet_PanglaoMouse, unique(unlist(GeneSet_PanglaoMouse)))
GeneSetCor(Ollaps_Allen)
ggsave(filename = paste0("Outputs/", date, "_GeneSet_PanglaoMouse_cor.png"))
```

Read the file with the average expression of each cluster

```{r}
# from file: "Cluster_investigation_3rdDataSet_prolijo.R"
date = 240725
id <- "all" #"4th" #"3rd" 
clusters_ctrl <- readRDS(paste0("Outputs/", date, "_Clusters_avg_exp_int_ctrl_", id, ".rds"))
#clusters_ctrl <- select(clusters_ctrl, names(clusters_ctrl)[!(names(clusters_ctrl) %in% c("19","18","17","16","15","12"))])
clusters_ctrl <- mutate_all(clusters_ctrl[-length(clusters_ctrl)], as.numeric) %>% as.matrix() #column 24 is "gene_name"
id <- "all"
date = 240725
```

### Plot Seurat Average Expression (just schematic)

```{r}
#pdf(paste0("Outputs/", date, "_AvgExpressionCtrl_hm", id, ".pdf"))
prueba <- clusters_ctrl[sample(nrow(clusters_ctrl),50),]
rownames(prueba) <- NULL
heatmap.2 (t(rbind(prueba[sample(nrow(prueba)),], prueba[sample(nrow(prueba)),])),
           # dendrogram control
           Rowv = FALSE,
           Colv= TRUE,
           col = "bluered",
           breaks = unique(c(seq(-1, -0.3, length=51), seq(-0.3, 0.3, length=101), seq(0.3, 1, length=51))),
           distfun = dist,
           trace = "none",
           hclustfun = hclust,
           dendrogram = "none",
           scale = "row")
#dev.off()
```

## Run GSVA analysis

#### CellMarker (Mouse)

```{r, include = F}
enrichCtrl <- gsva(clusters_ctrl, GeneSet_CellMarkerMouse)
```

```{r}
d_res <- hclust(dist(t(enrichCtrl)))
d_res <- cutree(d_res, k = 2)
# plot
## Parameters
d_order = c(1,2)
CellType_unique = c("Neuron", "Others")
# In the future I will put:
# CellType_unique = c("Neuron", "Others", "Microglia", "LC")
# So:
#viridis_pal(option = "D")(7)[-c(2,4,7)] #4 colors
colors_general <- c("#31688EFF", "#8FD744FF", "#35B779FF", "#440154FF")
## Data preparation
# annotate
PrepData_all <- PrepareDataH(enrichCtrl, d_res, d_order, CellType_unique, colors_general[c(1,2)])
saveRDS(PrepData_all, paste0("Outputs/",date, "_PrepData_CellMarkerMouse", id, ".rds"))
## Plot HM
#pdf(paste0("Outputs/", date, "_enrichCtrl_hm_GeneSet_general", id, ".pdf"))
PlotHM(enrichCtrl, PrepData_all)
#dev.off()
```

```{r, include = F}
enrichCtrl <- gsva(clusters_ctrl, GeneSet_CellMarkerMouse)
```

```{r}
d_res <- hclust(dist(t(enrichCtrl)))
d_res <- cutree(d_res, k = 3)
# plot
## Parameters
d_order = c(1,2,3)
CellType_unique = c("Neuron", "Others", "Microglia")
## Data preparation
# annotate
PrepData_all <- PrepareDataH(enrichCtrl, d_res, d_order, CellType_unique, colors_general[c(1,2,3)])
saveRDS(PrepData_all, paste0("Outputs/",date, "_PrepData_CellMarkerMouse", id, ".rds"))
## Plot HM
#pdf(paste0("Outputs/", date, "_enrichCtrl_hm_GeneSet_general", id, ".pdf"))
PlotHM(enrichCtrl, PrepData_all)
#dev.off()
```

#### CellMarker (Human)

```{r, include=F}
enrichCtrl_LC <- gsva(clusters_ctrl, GeneSet_CellMarkerHuman)
```

```{r}
d_res <- hclust(dist(t(enrichCtrl_LC)))
d_res <- cutree(d_res, k = 4)
# plot
## Parameters
d_order = c(1,2,3,4)
CellType_unique = c("Neuron", "Others", "Microglia", "LC")
## Data preparation
# annotate
PrepData <- PrepareDataH(enrichCtrl_LC, d_res, d_order, CellType_unique, colors_general)
saveRDS(PrepData, paste0("Outputs/",date, "_PrepData_GeneSet_CellMarkerHuman", id, ".rds"))
## Get legend
#pdf(paste0("Outputs/", date, "_enrichCtrl_legend_GeneSet_general_and_LC", id, ".pdf"))
#as.ggplot(get_legend(plot4legend(PrepData)))
#dev.off()
## Plot HM
#pdf(paste0("Outputs/", date, "_enrichCtrl_hm_GeneSet_general_and_LC", id, ".pdf"))
PlotHM(enrichCtrl_LC, PrepData)
#dev.off()
```

#### Panglao (Mouse)

```{r, include = F}
enrichCtrl_Allen <- gsva(clusters_ctrl, GeneSet_PanglaoMouse)
```

```{r}
d_res <- hclust(dist(t(enrichCtrl_Allen)))
d_res <- cutree(d_res, k = 5)
## Plot HeatMap of GSVA data
## Parameters
#dres <- c(1,2,3,4,5)
#d_order = c(1,4,5,3,2)
d_order = c(1,5,4,2,3)
CellType_unique = c("Not Ident", "LC", "Neigh or LC", "Neigh2 or LC", "Distant")
## Data preparation
#viridis_pal(option = "C")(7)[-c(1,6,7)]
colors_allen <- c("gray", "#5D01A6FF", "#9C179EFF", "#CC4678FF", "#ED7953FF")
color_table_allen <- data.frame(CellType_unique, colors = colors_allen)
PrepData <- PrepareDataH(enrichCtrl_Allen, d_res, d_order, CellType_unique,  colors_allen)
saveRDS(PrepData, paste0("Outputs/",date, "_PrepData_GeneSet_PanglaoMouse", id, ".rds"))
## Get legend
#pdf(paste0("Outputs/", date, "_enrichCtrl_legend_GeneSet_Allen", id, ".pdf"))
#as.ggplot(get_legend(plot4legend(PrepData))
#dev.off()
## Plot HM
#pdf(paste0("Outputs/", date, "_enrichCtrl_hm_GeneSet_Allen", id, ".pdf"))
PlotHM(enrichCtrl_Allen, PrepData)
#dev.off()
```

## For the 3rdDS


```{r}
# from file: "Cluster_investigation_3rdDataSet_prolijo.R"
date = 240725
id <- "3rd" 
clusters_ctrl <- readRDS(paste0("Outputs/", date, "_Clusters_avg_exp_int_ctrl_", id, ".rds"))
#clusters_ctrl <- select(clusters_ctrl, names(clusters_ctrl)[!(names(clusters_ctrl) %in% c("19","18","17","16","15","12"))])
clusters_ctrl <- mutate_all(clusters_ctrl[-length(clusters_ctrl)], as.numeric) %>% as.matrix() #column 24 is "gene_name"
id <- "3rd"
date = 240725
```

### Plot Seura tAverage Expression (just schematic)

```{r}
#pdf(paste0("Outputs/", date, "_AvgExpressionCtrl_hm", id, ".pdf"))
prueba <- clusters_ctrl[sample(nrow(clusters_ctrl),50),]
rownames(prueba) <- NULL
heatmap.2 (t(rbind(prueba[sample(nrow(prueba)),], prueba[sample(nrow(prueba)),])),
           # dendrogram control
           Rowv = FALSE,
           Colv= TRUE,
           col = "bluered",
           breaks = unique(c(seq(-1, -0.3, length=51), seq(-0.3, 0.3, length=101), seq(0.3, 1, length=51))),
           distfun = dist,
           trace = "none",
           hclustfun = hclust,
           dendrogram = "none",
           scale = "row")
#dev.off()
```

## Run GSVA analysis

#### CellMarker (Mouse)

```{r, include = F}
enrichCtrl <- gsva(clusters_ctrl, GeneSet_CellMarkerMouse)
```

```{r}
d_res <- hclust(dist(t(enrichCtrl)))
d_res <- cutree(d_res, k = 2)
# plot
## Parameters
d_order = c(1,2)
CellType_unique = c("Neuron", "Others")
# In the future I will put:
# CellType_unique = c("Neuron", "Others", "Microglia", "LC")
# So:
#viridis_pal(option = "D")(7)[-c(2,4,7)] #4 colors
colors_general <- c("#31688EFF", "#8FD744FF", "#35B779FF", "#440154FF")
## Data preparation
# annotate
PrepData_all <- PrepareDataH(enrichCtrl, d_res, d_order, CellType_unique, colors_general[c(1,2)])
saveRDS(PrepData_all, paste0("Outputs/",date, "_PrepData_CellMarkerMouse", id, ".rds"))
## Plot HM
#pdf(paste0("Outputs/", date, "_enrichCtrl_hm_GeneSet_general", id, ".pdf"))
PlotHM(enrichCtrl, PrepData_all)
#dev.off()
```

```{r, include = F}
enrichCtrl <- gsva(clusters_ctrl, GeneSet_CellMarkerMouse)
```

```{r}
d_res <- hclust(dist(t(enrichCtrl)))
d_res <- cutree(d_res, k = 3)
# plot
## Parameters
d_order = c(1,2,3)
CellType_unique = c("Neuron", "Others", "Microglia")
## Data preparation
# annotate
PrepData_all <- PrepareDataH(enrichCtrl, d_res, d_order, CellType_unique, colors_general[c(1,2,3)])
saveRDS(PrepData_all, paste0("Outputs/",date, "_PrepData_CellMarkerMouse", id, ".rds"))
## Plot HM
#pdf(paste0("Outputs/", date, "_enrichCtrl_hm_GeneSet_general", id, ".pdf"))
PlotHM(enrichCtrl, PrepData_all)
#dev.off()
```

#### CellMarker (Human)

```{r, include=F}
enrichCtrl_LC <- gsva(clusters_ctrl, GeneSet_CellMarkerHuman)
```

```{r}
d_res <- hclust(dist(t(enrichCtrl_LC)))
d_res <- cutree(d_res, k = 4)
# plot
## Parameters
d_order = c(1,2,3,4)
CellType_unique = c("Neuron", "Others", "Microglia", "LC")
## Data preparation
# annotate
PrepData <- PrepareDataH(enrichCtrl_LC, d_res, d_order, CellType_unique, colors_general)
saveRDS(PrepData, paste0("Outputs/",date, "_PrepData_GeneSet_CellMarkerHuman", id, ".rds"))
## Get legend
#pdf(paste0("Outputs/", date, "_enrichCtrl_legend_GeneSet_general_and_LC", id, ".pdf"))
#as.ggplot(get_legend(plot4legend(PrepData)))
#dev.off()
## Plot HM
#pdf(paste0("Outputs/", date, "_enrichCtrl_hm_GeneSet_general_and_LC", id, ".pdf"))
PlotHM(enrichCtrl_LC, PrepData)
#dev.off()
```

#### Panglao (Mouse)

```{r, include = F}
enrichCtrl_Allen <- gsva(clusters_ctrl, GeneSet_PanglaoMouse)
```

```{r}
d_res <- hclust(dist(t(enrichCtrl_Allen)))
d_res <- cutree(d_res, k = 5)
## Plot HeatMap of GSVA data
## Parameters
#dres <- c(1,2,3,4,5)
#d_order = c(1,4,5,3,2)
d_order = c(1,5,4,2,3)
CellType_unique = c("Not Ident", "LC", "Neigh or LC", "Neigh2 or LC", "Distant")
## Data preparation
#viridis_pal(option = "C")(7)[-c(1,6,7)]
colors_allen <- c("gray", "#5D01A6FF", "#9C179EFF", "#CC4678FF", "#ED7953FF")
color_table_allen <- data.frame(CellType_unique, colors = colors_allen)
PrepData <- PrepareDataH(enrichCtrl_Allen, d_res, d_order, CellType_unique,  colors_allen)
saveRDS(PrepData, paste0("Outputs/",date, "_PrepData_GeneSet_PanglaoMouse", id, ".rds"))
## Get legend
#pdf(paste0("Outputs/", date, "_enrichCtrl_legend_GeneSet_Allen", id, ".pdf"))
#as.ggplot(get_legend(plot4legend(PrepData))
#dev.off()
## Plot HM
#pdf(paste0("Outputs/", date, "_enrichCtrl_hm_GeneSet_Allen", id, ".pdf"))
PlotHM(enrichCtrl_Allen, PrepData)
#dev.off()
```

## For the 4thDS


```{r}
# from file: "Cluster_investigation_3rdDataSet_prolijo.R"
date = 240725
id <- "4th" 
clusters_ctrl <- readRDS(paste0("Outputs/", date, "_Clusters_avg_exp_int_ctrl_", id, ".rds"))
#clusters_ctrl <- select(clusters_ctrl, names(clusters_ctrl)[!(names(clusters_ctrl) %in% c("19","18","17","16","15","12"))])
clusters_ctrl <- mutate_all(clusters_ctrl[-length(clusters_ctrl)], as.numeric) %>% as.matrix() #column 24 is "gene_name"
id <-  "4th"
date = 240725
```

### Plot SeuratAverage Expression (just schematic)

```{r}
#pdf(paste0("Outputs/", date, "_AvgExpressionCtrl_hm", id, ".pdf"))
prueba <- clusters_ctrl[sample(nrow(clusters_ctrl),50),]
rownames(prueba) <- NULL
heatmap.2 (t(rbind(prueba[sample(nrow(prueba)),], prueba[sample(nrow(prueba)),])),
           # dendrogram control
           Rowv = FALSE,
           Colv= TRUE,
           col = "bluered",
           breaks = unique(c(seq(-1, -0.3, length=51), seq(-0.3, 0.3, length=101), seq(0.3, 1, length=51))),
           distfun = dist,
           trace = "none",
           hclustfun = hclust,
           dendrogram = "none",
           scale = "row")
#dev.off()
```

## Run GSVA analysis

#### CellMarker (Mouse)

```{r, include = F}
enrichCtrl <- gsva(clusters_ctrl, GeneSet_CellMarkerMouse)
```

```{r}
d_res <- hclust(dist(t(enrichCtrl)))
d_res <- cutree(d_res, k = 2)
# plot
## Parameters
d_order = c(1,2)
CellType_unique = c("Neuron", "Others")
# In the future I will put:
# CellType_unique = c("Neuron", "Others", "Microglia", "LC")
# So:
#viridis_pal(option = "D")(7)[-c(2,4,7)] #4 colors
colors_general <- c("#31688EFF", "#8FD744FF", "#35B779FF", "#440154FF")
## Data preparation
# annotate
PrepData_all <- PrepareDataH(enrichCtrl, d_res, d_order, CellType_unique, colors_general[c(1,2)])
saveRDS(PrepData_all, paste0("Outputs/",date, "_PrepData_CellMarkerMouse", id, ".rds"))
## Plot HM
#pdf(paste0("Outputs/", date, "_enrichCtrl_hm_GeneSet_general", id, ".pdf"))
PlotHM(enrichCtrl, PrepData_all)
#dev.off()
```

```{r, include = F}
enrichCtrl <- gsva(clusters_ctrl, GeneSet_CellMarkerMouse)
```

```{r}
d_res <- hclust(dist(t(enrichCtrl)))
d_res <- cutree(d_res, k = 3)
# plot
## Parameters
d_order = c(1,2,3)
CellType_unique = c("Neuron", "Others", "Microglia")
## Data preparation
# annotate
PrepData_all <- PrepareDataH(enrichCtrl, d_res, d_order, CellType_unique, colors_general[c(1,2,3)])
saveRDS(PrepData_all, paste0("Outputs/",date, "_PrepData_CellMarkerMouse", id, ".rds"))
## Plot HM
#pdf(paste0("Outputs/", date, "_enrichCtrl_hm_GeneSet_general", id, ".pdf"))
PlotHM(enrichCtrl, PrepData_all)
#dev.off()
```

#### CellMarker (Human)

```{r, include=F}
enrichCtrl_LC <- gsva(clusters_ctrl, GeneSet_CellMarkerHuman)
```

```{r}
d_res <- hclust(dist(t(enrichCtrl_LC)))
d_res <- cutree(d_res, k = 4)
# plot
## Parameters
d_order = c(1,2,3,4)
CellType_unique = c("Neuron", "Others", "Microglia", "LC")
## Data preparation
# annotate
PrepData <- PrepareDataH(enrichCtrl_LC, d_res, d_order, CellType_unique, colors_general)
saveRDS(PrepData, paste0("Outputs/",date, "_PrepData_GeneSet_CellMarkerHuman", id, ".rds"))
## Get legend
#pdf(paste0("Outputs/", date, "_enrichCtrl_legend_GeneSet_general_and_LC", id, ".pdf"))
#as.ggplot(get_legend(plot4legend(PrepData)))
#dev.off()
## Plot HM
#pdf(paste0("Outputs/", date, "_enrichCtrl_hm_GeneSet_general_and_LC", id, ".pdf"))
PlotHM(enrichCtrl_LC, PrepData)
#dev.off()
```

#### Panglao (Mouse)

```{r, include = F}
enrichCtrl_Allen <- gsva(clusters_ctrl, GeneSet_PanglaoMouse)
```

```{r}
d_res <- hclust(dist(t(enrichCtrl_Allen)))
d_res <- cutree(d_res, k = 5)
## Plot HeatMap of GSVA data
## Parameters
#dres <- c(1,2,3,4,5)
#d_order = c(1,4,5,3,2)
d_order = c(1,5,4,2,3)
CellType_unique = c("Not Ident", "LC", "Neigh or LC", "Neigh2 or LC", "Distant")
## Data preparation
#viridis_pal(option = "C")(7)[-c(1,6,7)]
colors_allen <- c("gray", "#5D01A6FF", "#9C179EFF", "#CC4678FF", "#ED7953FF")
color_table_allen <- data.frame(CellType_unique, colors = colors_allen)
PrepData <- PrepareDataH(enrichCtrl_Allen, d_res, d_order, CellType_unique,  colors_allen)
saveRDS(PrepData, paste0("Outputs/",date, "_PrepData_GeneSet_PanglaoMouse", id, ".rds"))
## Get legend
#pdf(paste0("Outputs/", date, "_enrichCtrl_legend_GeneSet_Allen", id, ".pdf"))
#as.ggplot(get_legend(plot4legend(PrepData))
#dev.off()
## Plot HM
#pdf(paste0("Outputs/", date, "_enrichCtrl_hm_GeneSet_Allen", id, ".pdf"))
PlotHM(enrichCtrl_Allen, PrepData)
#dev.off()
```


