---
title: "QC Analysis and Integration 1st DataSet SN"
author: "Camila Arcuschin; Ignacio Schor"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

```{r setup, include=T}
knitr::opts_chunk$set(message = FALSE, echo = TRUE, warning = FALSE)
knitr::opts_knit$set(root.dir = "/data_husihuilke/shared/SMB/carcu/Parkinsons_project/")
```

```{r importpckg, include=T}
library(Seurat)
library(tidyverse)
library(hdf5r)
library(pryr)
library(gridExtra)
library(grid)
library(ragg) # scaling plots
library(reticulate)
library(patchwork)
library(cowplot)
library(viridis)
library(future) # parallel processing
library(scran)
library(scDblFinder)
library(scater)
RhpcBLASctl::blas_set_num_threads(3)
plan(multisession, workers = 5)
options(future.globals.maxSize = 4000 * 1024^2)
#py_install("pandas")
wd = "/data_husihuilke/shared/SMB/carcu/Parkinsons_project/"
```

```{r Functions, include =T}
#### Functions ####

# Create Seurat Object and filter by UMIs.p.cell and ribosomal percentage
Seu_filter <-
  function(dataX,
           UMIs.p.cell_lth,
           UMIs.p.cell_hth,
           percent.rb_th,
           treatment) {
    # Implement some filter parameters to avoid working with a very big matrix
    # Gene must be expressed in at least 30 cell (of the 2452867 "cells")
    # Explore the data, rename the cells metadata
    GetAssayData(dataX)
    dataX[[]] %>% str()
    dataX@meta.data <-
      rename(dataX@meta.data, "nFeature_RNA" = "ActiveGenes.p.cell")
    dataX@meta.data <-
      mutate(dataX@meta.data, "UMIs.p.cell" = nCount_RNA)
    #Add MetaData
    is.ribo <-
      lapply(c("Rpl", "Rps"), function(x)
        grepl(x, rownames(dataX))) %>% unlist()
    ribo_genes <-
      lapply(c("Rpl", "Rps"), function(x)
        grep(x, rownames(dataX), value = T)) %>% unlist()
    ribo_umis <- GetAssayData(dataX)[ribo_genes,] %>% colSums()
    dataX[["percent.rb"]] <- 100 * ribo_umis / dataX$UMIs.p.cell
    dataX[["percent.mt"]] <-
      PercentageFeatureSet(dataX, pattern = "Mt")
    # What would happen if I only kept the cell with more than 2000 UMIs (aprox remove the 7% of the distribution)
    dataX <-
      subset(
        dataX,
        subset = UMIs.p.cell > UMIs.p.cell_lth &
          UMIs.p.cell < UMIs.p.cell_hth & percent.rb < percent.rb_th
      )
    return(dataX)
  }

get_legend <- function(myggplot) {
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x)
    x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

# Log Normalize Seurat and select the 2000 most variable genes
norm_seurat <- function(all_samples) {
  all_samples <-
    NormalizeData(all_samples,
                  normalization.method = "LogNormalize",
                  scale.factor = 10000)
  all_samples <-
    FindVariableFeatures(
      all_samples,
      selection.method = "vst",
      nfeatures = 2000,
      loess.span = 0.6
    )
}

# Run PCA, knn and UMAP
UMAP_maker <- function(obj, dims, k.param) {
  ScaleData(object = obj) %>%
    RunPCA(npcs = dims) %>%
    FindNeighbors(dims = 1:dims, k.param = k.param) %>% #Esta linea no har?a falta, igual creo que no cambiar?a nada porque siempre lo estoy pisando.
    RunUMAP(dims = 1:dims)
} # The reason why I have FindNeighbours here is because, if I want to replot UMAP whithout changing the cluster already assigned, I could do it

preprocess_RPCA <- function(obj.list, features) {
  obj.list <- ScaleData(object = obj.list, features = features) %>%
    RunPCA(features = features)
}

fisher <- function(x) {
  matriz <- matrix(c(x[1], x[2], x[3], x[4]), nrow = 2)
  f_test <- fisher.test(matriz)
  data_fisher <-
    c(enrich = f_test[["estimate"]][["odds ratio"]],
      p_val = f_test[["p.value"]],
      conf_int = f_test[["conf.int"]])
  return(data_fisher)
}

create_random <- function(To, p_E, p_S, p_N){
  vec <- sample(c("EGFP", "SNCA", "EGFP_SNCA", "Neith"), To,  replace = T, prob = c(p_E-p_E*p_S, p_S-p_E*p_S, p_E*p_S, p_N))
  table(vec)
}

EGFP_SNCA <- function(df) {
  nrow(filter(df, `EGFP` == 1 & `hSNCA` == 1))
}

EGFP <- function(df) {
  nrow(filter(df, `EGFP` == 1 & `hSNCA` == 0))
}

SNCA <- function(df) {
  nrow(filter(df, `EGFP` == 0 & `hSNCA` == 1))
}

Neith <- function(df) {
  nrow(filter(df, `EGFP` == 0 & `hSNCA` == 0))
}

label2analyze <- function(FishTest){
  vect <- c()
  for (i in 1:length(clusters2analyze)){
    if (i != 3){
      vect[i] <- paste0("Cluster ", clusters2analyze[i], " = ", round(FishTest["enrich", clusters2analyze[i]], 2), " (p-adj = ", format(FishTest["FDR", clusters2analyze[i]], digits=3), ") ; ")
    } else {
      vect[i] <- paste0("Cluster ", clusters2analyze[i], " = ", round(FishTest["enrich", clusters2analyze[i]], 2), " (p-adj = ", format(FishTest["FDR", clusters2analyze[i]], digits=3), ") ;\n")
    }
  }
  return(vect)
}

# Plotting
my_theme <- theme_classic() +
  theme(
    plot.title = element_text(size = 25, face = "bold"),
    axis.text.x = element_text(colour = "black", size = 14),
    axis.text.y = element_text(colour = "black", size = 14),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    axis.title.x = element_text(size = 15),
    strip.text.x = element_text(size = 15)
  )

QC_plots <- function(SeuObj, extra_text, n_color) {
  Dist_QC <-
    VlnPlot(
      SeuObj,
      features = c(
        "ActiveGenes.p.cell",
        "UMIs.p.cell",
        "percent.mt",
        "percent.rb"
      ),
      ncol = 4,
      pt.size = 0,
      cols = n_color
    ) &
    theme(legend.position = "none",
          axis.title = element_text(size = 6),
          axis.text = element_text(size = 5),
          plot.title = element_text(size = 8, face = "bold"))
  stth <-     theme(
    legend.position = "none",
    axis.title = element_text(size = 6),
    axis.text = element_text(size = 5),
    plot.title = element_text(size = 8, face = "bold")
  )
  plot1 <-
    FeatureScatter(
      SeuObj,
      feature1 = "UMIs.p.cell",
      feature2 = "percent.mt",
      pt.size = 0.5,
      plot.cor =  F,
      cells = sample(Cells(SeuObj)),
      cols = n_color
    ) + stth
  plot2 <-
    FeatureScatter(
      SeuObj,
      feature1 = "UMIs.p.cell",
      feature2 = "ActiveGenes.p.cell",
      pt.size = 0.5,
      plot.cor =  F,
      cells = sample(Cells(SeuObj)),
      cols = n_color
    ) + stth
  plot3 <-
    FeatureScatter(
      SeuObj,
      feature1 = "ActiveGenes.p.cell",
      feature2 = "percent.rb",
      pt.size = 0.5,
      plot.cor =  F,
      cells = sample(Cells(SeuObj)),
      cols = n_color
    )+ stth
  plot4 <-
    FeatureScatter(
      SeuObj,
      feature1 = "UMIs.p.cell",
      feature2 = "percent.rb",
      pt.size = 0.5,
      plot.cor =  F,
      cells = sample(Cells(SeuObj)),
      cols = n_color
    ) + stth
  Scat_QC <-
    plot_grid(plot1, plot2, plot3, plot4, nrow = 1)
  QC_plot <- (Dist_QC / Scat_QC) +
    plot_annotation(
      title = paste0(
        extra_text,
        "Quality Control Metrics for > ",
        UMIs.p.cell_lowthreshold,
        " <",
        UMIs.p.cell_highthreshold,
        " UMI < ",
        percent.rb_threshold,
        " per.rb Matrix (",
        dim(SeuObj)[2],
        " Cells)"
      ),
      subtitle = "each number is a cluster",
      theme = theme(
        axis.text = element_text(size = 5),
        axis.title = element_text(size = 6),
        plot.title = element_text(size = 8, face = "bold")
      )
    )
  return(QC_plot)
}

UMAP_by_clust <-
  function(SeuObj,
           ClusterBreaks,
           ClusterLabels,
           pt_size,
           title_text,
           subtitle_text,
           color_a,
           color_b,
           do_label,
           n_color) {
    (
      UMAPa <-
        DimPlot(
          SeuObj,
          reduction = "umap",
          label = do_label,
          pt.size = pt_size
        ) +
         scale_colour_manual(values = n_color, breaks = ClusterBreaks, labels = ClusterLabels) +
        labs(
          title = title_text,
          subtitle = subtitle_text
          # color = color_a
        ) +
        theme(
          axis.title = element_text(size = 6),
          legend.title = element_text(size = 6),
          legend.text = element_text(size = 6),
          legend.position = "bottom",
          plot.title = element_text(size = 10, face = "bold"), axis.text = element_text(size = 0), axis.ticks = element_line(size = 0)
        ) +
        guides(color = guide_legend(
          title.position="top",
          override.aes = list(size = 1), nrow = 3
        ))
    )
    (
      UMAPb <-
        DimPlot(
          SeuObj,
          reduction = "umap",
          label = F,
          pt.size = pt_size,
          group.by = "Treat",
          cells = sample(Cells(SeuObj))
        ) +
        labs(
          title = "",
          subtitle = "",
          color = color_b
        ) +
        theme(
          legend.position = "right",
          axis.title = element_text(size = 6),
          legend.title = element_text(size = 6),
          legend.text = element_text(size = 6),
          plot.title = element_text(size = 10, face = "bold"), axis.text = element_text(size = 0), axis.ticks = element_line(size = 0)
        ) +
        guides(color = guide_legend(
          override.aes = list(size = 1), nrow = 2
        ))
    )
    return(list("UMAPa" = UMAPa, "UMAPb" = UMAPb))
  }
# UMAP_by_clust <-
#   function(SeuObj,
#            ClusterBreaks,
#            ClusterLabels,
#            pt_size,
#            title_text,
#            subtitle_text,
#            color_a,
#            color_b,
#            do_label) {
#     (
#       UMAPa <-
#         DimPlot(
#           SeuObj,
#           reduction = "umap",
#           label = do_label,
#           pt.size = pt_size
#         ) +
#         scale_colour_discrete(breaks = ClusterBreaks,
#                               labels = ClusterLabels) +
#         labs(
#           title = title_text,
#           subtitle = subtitle_text,
#           color = color_a
#         ) +
#         theme(
#           axis.title = element_text(size = 6),
#           legend.title = element_text(size = 6),
#           legend.text = element_text(size = 6),
#           legend.position = "bottom",
#           plot.title = element_text(size = 10, face = "bold")
#         ) +
#         guides(color = guide_legend(
#           title.position="top",
#           override.aes = list(size = 1), nrow = 3
#         ))
#     )
#     (
#       UMAPb <-
#         DimPlot(
#           SeuObj,
#           reduction = "umap",
#           label = F,
#           pt.size = pt_size,
#           group.by = "Treat",
#           cells = sample(Cells(SeuObj))
#         ) +
#         labs(
#           title = "",
#           subtitle = "",
#           color = color_b
#         ) +
#         theme(
#           legend.position = "right",
#           axis.title = element_text(size = 6),
#           legend.title = element_text(size = 6),
#           legend.text = element_text(size = 6),
#           plot.title = element_text(size = 10, face = "bold")
#         ) +
#         guides(color = guide_legend(
#           override.aes = list(size = 1), nrow = 2
#         ))
#     )
#     return(list("UMAPa" = UMAPa, "UMAPb" = UMAPb))
#   }

insert_minor <- function(major_labs, n_minor) {
  labs <-
    c(sapply(major_labs, function(x)
      c(x, rep("", 4))))
  labs[1:(length(labs) - n_minor)]
}

UMAPs_for_grid <- function(SeuObj, subtitle_text) {
  Idents(SeuObj) <- "Genes"
  UMAP4 <-
    DimPlot(
      SeuObj,
      reduction = "umap",
      order = T,
      label = F,
      pt.size = 1,
      label.size = 8,
      ncol = 1,
      cols = c("grey", "#40a1e4", "#e48340", "#4c0e97")
    ) +
    labs(title = "", subtitle = "EGFP-KASH+ hSNCA-A53T+ Cells") + theme(legend.position = "top") +
    guides(color = guide_legend(override.aes = list(size = 4), nrow = 1))
  legend <- get_legend(UMAP4)
  UMAP4 <- UMAP4 + theme(legend.position = "none")
  Idents(SeuObj) <- "seurat_clusters"
  UMAPa <-
    DimPlot(SeuObj,
            reduction = "umap",
            pt.size = 0.7,
            label = F,
            cols = colors_n) + theme(legend.position = "none")
  UMAPa_edit <-
    LabelClusters(UMAPa,
                  id = "ident",
                  size = 4,
                  repel = F) + labs(title = "", subtitle = subtitle_text) +
    theme(title = element_text(size = 15))
  return(list(
    "UMAP4_split" = UMAP4,
    "UMAPa_edit" = UMAPa_edit,
    "UMAP4_legend" = legend
  ))
}

FeaturePlotMultiple2 <-
  function(obj, feature, color, ord, alf, minlim) {
    # # the minimal and maximal of the value to make the legend scale the same.
    # minimal<- min(obj[['RNA']]@data[feature, ])
    # maximal<- max(obj[['RNA']]@data[feature, ])
    ps <- vector("list", length(feature))
    for (i in 1:length(feature)) {
      if (ord) {
        p <-
          FeaturePlot(
            obj,
            features = feature[i],
            min.cutoff = minlim,
            max.cutoff = "q99",
            pt.size = 0.00001,
            cols = c(alpha("lightgrey", alf), alpha(color[i], alf)),
            order = T
          )
      } else {
        p <-
          FeaturePlot(
            obj,
            features = feature[i],
            min.cutoff = minlim,
            max.cutoff = "q99",
            pt.size = 0.00001,
            cols = c(alpha("lightgrey", alf), alpha(color[i], alf)),
            cells = sample(Cells(obj))
          )
      }
      #ggtitle(feature[i]) + theme_void() +
      p <-
        p + theme(
          text = element_text(size = 6),
          axis.text = element_text(size = 2.7),
          plot.title = element_text(size = 10, face = "bold"),
          legend.position = "none"
        )
      plot_name <- feature[i]
      ps[[i]] <- p
      names(ps)[[i]] <- plot_name
    }
    return(ps)
  }
```

```{r ImportInput, include = T, echo = T}
## Date ##

 date = 231122 # La corrida que va a sacar intra-samples doublets con el dbfinddoublets filter 4thDS
 numDS = "1st"
 equiv_values_Treat = c("PD", "Ctrl", "PD", "Ctrl") # 4thDS
 equiv_values_Age = c("2mo" , "1mo", "1mo", "2mo") # 4thDS
 colors_hash = c("#00BFC4", "#F8766D", "#C77CFF", "#D39200")
 clusters2analyze = c(0,1,2,3,4,5,6,7,8,9,10) %>% as.character()
 text4stackbar = c(0,1,2,4,5,6,7) %>% as.character()
 text4stackbarEGFP = c(0,6,8,9) %>% as.character()
 FilePath = "Data/230601/SCC0117_Briese_SNc_custom/outs/" # la corrida del 1stDS
 datadate = "231122 SNc"
 cluster_map = c(9,8,7,3,6,5,10) %>% as.character()

Cells_SN_df <- readRDS("Outputs/Cre_Cells_SN_df.rds")
```

**Extra data to take into account:**

From 10x CellRanger:

-   "Our internal tests have shown the recovery of Single Cell
    populations between whole cells and nuclei is comparable. In some
    cases, there are advantages to using single nuclei for RNA-seq,
    including reducing biases in cell populations and this is explained
    here.

-   "Please also read the technote on resolving cell types as a function
    of Read Depth and Cell Number here"

-   HTO Analysis:

    -   "If you also used antibody-based hashtag oligos (HTOs) for
        sample multiplexing, you may use the [multi pipeline in Cell
        Ranger
        6+](https://kb.10xgenomics.com/hc/en-us/articles/4407386498957-I-used-antibody-tags-for-cell-surface-protein-capture-and-cell-hashing-with-Single-Cell-3-chemistry-How-can-I-use-Cell-Ranger-to-analyze-my-data-)

## Data

Data used belong to the download from the `r datadate`. Reads have been
mapped to the mouse transcriptome plus transgenes incorporated:
KASH-EGFP and aSyn-A53T. Cells are pre-filetered by total UMI counts
(see the web_summary.html on the same folder).

```{r ImportInputReal, include = T}
## File Path/Name ####
FileName = "filtered_feature_bc_matrix.h5"
```

```{r include = T}
print("loading data") # Load the data and convert it into a Seurat Object
data <-
  Read10X_h5(paste0(FilePath, FileName),
             use.names = TRUE,
             unique.features = TRUE)
dim(data$`Gene Expression`) # 32285 7379 (date = 221102) #32285 22047 (date = 230522)
```

## Demultiplexing cells

The Seurat HTODemux function for demultiplexing HashTag antibody
associated Oligonucleotide (HTO) works by generating k clusters (with
k-medoids) of the cells based on the normalized HTO expression values
(being k = number of unique HTOs + 1 for the Negative group). Then, it
calculates the average expression values per cluster and assign the
Negative class to the cluster with the lowest expression. Within the
Negative cluster, it fits a Negative Binomial distribution for each HTO
and assign the 0.99 quantile of that distribution as a threshold for the
positivity or negativity of each HTO in each cell. Cells which are
positive for more than 1 HTO are classified as doublets.

```{r Demultiplex, include = T}
SeuObjRaw <-
  CreateSeuratObject(
    counts = data$`Gene Expression`,
    project = "FACSHash",
    min.cells = 0,
    min.features = 0
  ) # Create Seurat Object
SeuObjRaw[["HTO"]] <-
  CreateAssayObject(counts = data$`Antibody Capture`) # Add HTO assay
SeuObjRaw <-
  NormalizeData(SeuObjRaw, assay = "HTO", normalization.method = "CLR") # Normalize data
SeuObjRaw <-
  HTODemux(SeuObjRaw, assay = "HTO", positive.quantile = 0.99) # Demultiplex HTO according to HTO expression. I tried with kfunc =  "kmeans" but results doesn't chamge tht much and I lost 10 more cells
rm(data)
```

```{r include= T}
gc()
```

### Distribution of the summed HTO expression of each sample, according to hash.ID classification:

The line in red shows a threshold for the intra-sample doublets defined
by expression.

```{r RidgePlot_HTO_class}
RidgePlot_HTO_class <-
  RidgePlot(
    SeuObjRaw,
    assay = "HTO",
    features = "nCount_HTO" ,
    log = TRUE,
    ncol = 1
  )# Distribution of counts of negative, singlet or doublet classification.
RidgePlot_HTO_class <-
  RidgePlot_HTO_class + geom_vline(xintercept = 5000, colour = "red")
RidgePlot_HTO_class
```

### Distribution of each HTO raw expression:

```{r RidgePlot_HTO_expr, fig.asp=1.5, out.height= '25%'}
feat <- rownames(GetAssayData(SeuObjRaw[["HTO"]]))
Idents(SeuObjRaw) <- "all_cells"
RidgePlot_HTO_expr <-
  RidgePlot(
    SeuObjRaw,
    slot = "counts",
    assay = "HTO",
    features = feat,
    log = T,
    ncol = 1
  )# Distribution of counts of negative, singlet or doublet classification.
RidgePlot_HTO_expr <-
  RidgePlot_HTO_expr & geom_vline(xintercept = 800, colour = "red")
RidgePlot_HTO_expr
```

### Distribution of the summed HTO expression of each cell:

```{r RidgePlot_HTO_tgth, out.height= '25%', fig.asp=0.5}
RidgePlot_HTO_tgth <-
  RidgePlot(
    SeuObjRaw,
    assay = "HTO",
    features = "nCount_HTO" ,
    log = TRUE,
    ncol = 1
  )# Distribution of counts of negative, singlet or doublet classification.
RidgePlot_HTO_tgth <-
  RidgePlot_HTO_tgth + geom_vline(xintercept = 5000, colour = "red")
RidgePlot_HTO_tgth
```

### Distribution of the summed RNA expression of each cell type, according to being doublets, singlets or negatives:

```{r VlnPlot_HTO_global}
Idents(SeuObjRaw) <- "HTO_classification.global"
VlnPlot_HTO_global <-
  VlnPlot(SeuObjRaw,
          features = "nCount_RNA",
          pt.size = 0,
          log = TRUE) # Distribution of counts of negative, singlet or doublet classification.
VlnPlot_HTO_global
```

#### How many cells corresponds to each category?

```{r HTO_classification, include = T}
HTO_classification <-
  c(
    "singlet" = length(
      which(SeuObjRaw@meta.data$"HTO_classification.global" == "Singlet")
    ),
    "doublet" = length(
      which(SeuObjRaw@meta.data$"HTO_classification.global" == "Doublet")
    ),
    "negative" = length(
      which(
        SeuObjRaw@meta.data$"HTO_classification.global" == "Negative"
      )
    )
  )
print(HTO_classification)
```

## Discard negative -low counts- cells and analyze cells clustering by HTOs expression

Heatmap is done with HTO scaled expression

```{r HTOHeatmapAndTSNE}
SeuRaw_noNeg <- subset(SeuObjRaw, idents = "Negative", invert = TRUE)
rm(SeuObjRaw)

# Analize cells clustering by HTOs expression
DefaultAssay(SeuRaw_noNeg) <- "HTO"
SeuRaw_noNeg <-
  ScaleData(SeuRaw_noNeg,
            features = rownames(SeuRaw_noNeg),
            verbose = FALSE)
SeuRaw_noNeg <-
  RunPCA(
    SeuRaw_noNeg,
    features = rownames(SeuRaw_noNeg),
    approx = F,
    ndims.print = 1:2
  ) # RunPCA by default uses an approximate algorithm (irlba) for computing the PCs.
SeuRaw_noNeg <- RunTSNE(SeuRaw_noNeg, dims = 1:2, perplexity = 100)
tSNEPlot_HTO <-
  DimPlot(SeuRaw_noNeg, group.by = "hash.ID") # Plot UMAP, doublets colored in pink
tSNEPlot_HTO
HTOHeatmap <-
  HTOHeatmap(SeuRaw_noNeg,
             assay = "HTO",
             ncells = 5000,
             raster = F) # Plot HeatMap or each HTO level by cell in singlets and doublets
HTOHeatmap
```

```{r include =T}
gc()
```

#### Calculate the proportion of intra-sample doublets:

Given the fact that all the library preparation was done pooling
together cells from all samples with the HTO previously added, the
probability of a sample to match another sample is $1/(2^n)$, in which
$n$ is the number of samples. That gives a probability of $1/16$, in
which from this 16 cases only 4 corresponds to a self match (an
intra-sample doublet) and 12 to a non-self match (an inter-sample
doublet). This means there are 12 cases detected by the different HTO
expression in the same "cell/nuclei" but 4 cases undetected by this
method, that gives $1/3$ of the detected doublet cases ("known
doublets"). If there are `r HTO_classification["doublet"]` known
doublets, I would expect to have approx
`r HTO_classification["doublet"]`/3 =
`r HTO_classification["doublet"]/3`

```{r ExpressionDoublets, include=T}
possibleDoublet <-
  data.frame(hash.ID = as.character(SeuRaw_noNeg@meta.data$hash.ID)) %>% mutate(hash.ID.changed = if_else(SeuRaw_noNeg@meta.data$nCount_HTO > 5000, "Doublet", hash.ID))
rownames(possibleDoublet) <- NULL
SeuRaw_noNeg@meta.data$possibleDoublet <-
  possibleDoublet$hash.ID.changed %>% as.factor()
```

HeatMaps of each HTO expression only for cells assigned as doublets, including intra- and inter-sample doublets

```{r HTOHeatmap_known_and_expression_doublets, include=T}
#HeatMaps of each HTO expression only for cells assigned as doublets, including intra- and inter-sample doublets
HTOHeatmap_doublets <-
  HTOHeatmap(
    subset(SeuRaw_noNeg, subset = possibleDoublet == "Doublet") ,
    assay = "HTO",
    ncells = 5000,
    raster = F
  ) #, ncells = 5000)
HTOHeatmap_doublets
```

#### How many cells are intra-sample doublets identified by HTO expression?

There are
`r dim(subset(SeuRaw_noNeg, subset = possibleDoublet == "Doublet" & hash.ID != "Doublet"))[2]`
cells


Plot the HeatMaps of each HTO expression only for cells assigned as Doublets, intra- and inter-sample, but with the hash.ID classification
```{r HTOHeatmap_doublets_classification, include=T}
# Plot the HeatMaps of each HTO expression only for cells assigned as Doublets, intra- and inter-sample, but with the hash.ID classification
Idents(SeuRaw_noNeg) <- "hash.ID"
HTOHeatmap_doublets <-
  DoHeatmap(
    subset(SeuRaw_noNeg, subset = possibleDoublet == "Doublet") ,
    slot = "scale.data",
    assay = "HTO",
    features = feat
  ) #, ncells = 5000)
HTOHeatmap_doublets
#I can see graphically that approx 1/3 of the inter-treatment doublets could be intra-treatment doublets
```

## Apply the *scDblFinder* method to find intra-sample doublets by guilty association with inter-sample doublets

The method do the following:

-   Uses the PCA space to get the *k* nearest neighbors of each cell
    that are known doublets.

-   Also it pre-classifies cells by its more expressed HTO and estimate
    the proportion of cells in this categories.

-   Then it uses the number of *k* nearest -known doublets- neighbors
    that belong to the cell category and calculate a score of a
    "proportion of doublets" for each cell. As higher the proportion,
    higher the probability of that cell for being also a doublet.

-   A threshold of that score is determined to get all the doublets and
    the known doublets are excluded to identified the "predicted"
    intra-sample doublets.

    For more info see
    [here](http://bioconductor.org/books/3.15/OSCA.advanced/doublet-detection.html)
    or
    [here](http://www.bioconductor.org/packages/release/bioc/vignettes/scDblFinder/inst/doc/recoverDoublets.html)

```{r dbFindDoublet, include = T}
# DefaultAssay(SeuRaw_noNeg) <- "RNA"
# #SeuRaw_noNeg <- norm_seurat(SeuRaw_noNeg)
#   SeuRaw_noNeg <-
#     NormalizeData(SeuRaw_noNeg,
#                   normalization.method = "LogNormalize",
#                   scale.factor = 10000)
#   SeuRaw_noNeg <-
#     FindVariableFeatures(
#       SeuRaw_noNeg,
#       selection.method = "vst",
#       nfeatures = 4000,
#       loess.span = 0.6
#     )
# SeuRaw_noNeg <- ScaleData(SeuRaw_noNeg, verbose = FALSE)
# SeuRaw_noNeg <-
#   RunPCA(SeuRaw_noNeg, approx = T) #RunPCA by default uses an approximate algorithm (irlba) for computing the PCs.
# SeuRaw_noNeg <- RunTSNE(SeuRaw_noNeg, dims = 1:4, perplexity = 100)
# SeuRaw_noNeg <- RunUMAP(SeuRaw_noNeg, dims = 1:50)

UMIs.p.cell_lowthreshold = 0
UMIs.p.cell_highthreshold = Inf
percent.rb_threshold = Inf

DefaultAssay(SeuRaw_noNeg) <- "RNA"
SeuRaw_noNeg <-
  Seu_filter(
    SeuRaw_noNeg,
    UMIs.p.cell_lth = UMIs.p.cell_lowthreshold,
    UMIs.p.cell_hth = UMIs.p.cell_highthreshold,
    percent.rb_th = percent.rb_threshold,
    treatment = "Gene Expression"
  )
SeuRaw_noNeg <- norm_seurat(SeuRaw_noNeg)
SeuRaw_noNeg <- FindVariableFeatures(SeuRaw_noNeg)
SeuRaw_noNeg <-
  UMAP_maker(SeuRaw_noNeg, dims = 50, k.param = 10)
SeuRaw_noNeg <-
  RunTSNE(SeuRaw_noNeg,
          dims = 1:4,
          perplexity = 100)
SeuRaw_noNeg <-
  FindNeighbors(
    SeuRaw_noNeg,
    dims = 50,
    k.param = 10,
    compute.SNN = F
  )
SeuRaw_noNeg <-
  FindClusters(
    SeuRaw_noNeg,
    resolution = 0.05,
    n.start = 100,
    n.iter = 10
  )

sce.hash <-
  as.SingleCellExperiment(DietSeurat(SeuRaw_noNeg, graphs = c("pca", "tsne", "umap")))
# Recovering the intra-sample doublets:
colData(sce.hash)$Doublet <- colData(sce.hash)$hash.ID == "Doublet"
colData(sce.hash)$Doublet2 <-
  colData(sce.hash)$hash.ID != "Doublet" &
  colData(sce.hash)$possibleDoublet == "Doublet"
hashed.doublets <-
  recoverDoublets(
    sce.hash,
    use.dimred = "PCA",
    doublets = sce.hash$Doublet,
    samples = table(sce.hash$HTO_maxID),
    k = 100
  )
hashed.doublets
set.seed(1000101001)
sce.hash$proportion <- hashed.doublets$proportion
sce.hash$predicted <- hashed.doublets$predicted
```

### UMAP comparing the known, predicted and expression doublets

```{r UMAPSamplePredicted, fig.height= 7}
p0<-plotUMAP(sce.hash, colour_by = "seurat_clusters") + ggtitle("Clusters")
p0$layers[[1]]$aes_params$size <- 0.5
p1 <- plotUMAP(sce.hash, colour_by = "ActiveGenes.p.cell") + ggtitle("Active Genes")
p1$layers[[1]]$aes_params$size <- 0.5
p2<-plotUMAP(sce.hash, colour_by = "proportion") + ggtitle("Doublet proportions")
p2$layers[[1]]$aes_params$size <- 0.5
p3 <- plotUMAP(sce.hash, colour_by = "Doublet") + ggtitle("Known doublets")
p3$layers[[1]]$aes_params$size <- 0.5

UMAPSamplePredicted <- plot_grid(
    p0,
    p1,
    p2,
    p3,
  ggcells(sce.hash[, order(sce.hash$predicted)]) +
    geom_point(aes(
      x = UMAP.1, y = UMAP.2, colour = predicted
    ), size = 0.5) +
    scale_color_manual(values = c("grey", "red")) +
    theme_bw() +
    ggtitle("Predicted intra-sample doublets"),
  ggcells(sce.hash[, order(sce.hash$Doublet2)]) +
    geom_point(aes(
      x = UMAP.1, y = UMAP.2, colour = Doublet2
    ), size = 0.5) +
    scale_color_manual(values = c("grey", "blue")) +
    theme_bw() +
    ggtitle("Expression intra-sample doublets"),
  ncol = 2
)
UMAPSamplePredicted
```

Two cells from future dopaminergic cluster are lost here by intra or inter -sample filter
```{r}
sce.hash$CreCells <- colnames(sce.hash) %in% Cells_SN_df$Cells
which(sce.hash$CreCells)
Cells_SN_df$isInFinalDS <- ifelse(Cells_SN_df$isInFinalDS, "final", "no_final")
Cells_SN_df[match(colnames(sce.hash), Cells_SN_df$Cells, nomatch = 0)[match(colnames(sce.hash), Cells_SN_df$Cells, nomatch = 0) != 0],]
sce.hash$CreCells[sce.hash$CreCells] <- Cells_SN_df$isInFinalDS

plot_grid(
  p3,
ggcells(sce.hash[, order(sce.hash$predicted)]) +
    geom_point(aes(
      x = UMAP.1, y = UMAP.2, colour = predicted
    ), size = 0.5) +
    scale_color_manual(values = c("grey", "red")) +
    theme_bw() + theme(legend.position = "bottom") + 
    ggtitle("Predicted intra-sample doublets"),
ggcells(sce.hash[, order(sce.hash$CreCells)]) +
    geom_point(aes(
      x = UMAP.1, y = UMAP.2, colour = CreCells
    ), size = 0.5) +
    scale_color_manual(values = c("grey", "green", "purple")) +
    theme_bw() + theme(legend.position = "bottom") + 
    ggtitle("Cre Status"), ncol = 2, nrow = 2)

sce.hash$predicted[sce.hash$CreCells == "no_final"]
# DimPlot(SeuInt, cells.highlight = Cells_SN_df$Cells, label = T)
```

### tSNE comparing the known, predicted and expression doublets

```{r tSNESamplePredicted,  out.width= '150%', out.height= '75%', fig.asp=0.7}
tSNESamplePredicted <- plot_grid(
  plotTSNE(sce.hash, colour_by = "proportion") + ggtitle("Doublet proportions"),
  plotTSNE(sce.hash, colour_by = "Doublet") + ggtitle("Known doublets"),
  ggcells(sce.hash[, order(sce.hash$predicted)]) +
    geom_point(aes(
      x = TSNE.1, y = TSNE.2, colour = predicted
    )) +
    scale_color_manual(values = c("grey", "red")) +
    theme_bw() +
    ggtitle("Predicted intra-sample doublets"),
  ggcells(sce.hash[, order(sce.hash$Doublet2)]) +
    geom_point(aes(
      x = TSNE.1, y = TSNE.2, colour = Doublet2
    )) +
    scale_color_manual(values = c("grey", "blue")) +
    theme_bw() +
    ggtitle("Expression intra-sample doublets"),
  ncol = 2
)
tSNESamplePredicted
```

### "Proportion of doublets" score for each state: predicted, known and singlet

```{r hashed.doublets_vln}
state <-
  ifelse(
    hashed.doublets$predicted,
    "predicted",
    ifelse(hashed.doublets$known, "known", "singlet")
  )
colData(sce.hash) <- cbind(colData(sce.hash), state)
ggplot(as.data.frame(hashed.doublets)) + geom_violin(aes(x = state, y =
                                                           proportion))
```

#### How many *scDblFinder* intra-sample doublets belong to each of the categories obtained by intra-sample doublets found by HTO expression?

```{r expressionVspredicted, include=TRUE}
table(sce.hash$possibleDoublet, sce.hash$predicted)
  #              FALSE TRUE
  # Doublet       7326  656
  # SNc-Ctrl-1mo  3136  285
  # SNc-Ctrl-2mo  2505  289
  # SNc-PD-1mo    2621  522
  # SNc-PD-2mo    3830  512
```

#### How many doublets are ...:

-   predicted intra-sample, but not known inter-sample and also
    expression intra- and inter-sample:
    `r length(which(colData(sce.hash)$possibleDoublet == "Doublet" & colData(sce.hash)$hash.ID != "Doublet" & colData(sce.hash)$predicted == T))`

-   predicted intra-sample, but not known inter-sample:
    `r length(which(colData(sce.hash)$hash.ID != "Doublet" & colData(sce.hash)$predicted == T))`

-   predicted intra-sample and also expression intra- and inter-sample:
    `r length(which(colData(sce.hash)$possibleDoublet == "Doublet" & colData(sce.hash)$predicted == T))`

-   predicted intra-sample:
    `r length(which(colData(sce.hash)$predicted == T))`

-   expression intra- and -inter sample:
    `r length(which(colData(sce.hash)$possibleDoublet == "Doublet"))`

-   known inter-sample:
    `r length(which(colData(sce.hash)$hash.ID == "Doublet"))`

#### How many cells for each singlet, inter-sample (known) and intra-sample (predicted)

```{r totalDoublets, include = T}
print(table(sce.hash$state))
```

```{r save_sce.hash_object, include = FALSE}
## Save the sce Object
#saveRDS(sce.hash, paste0("Outputs/", date, "_sce.hash_", numDS, "DataSet.rds"))
```

```{r sce2Seurat, include = T}
## sce to Seurat
SeuRaw_noNeg <-
  as.Seurat(sce.hash, counts = "counts", data = "logcounts") #no deja que se agregue automaticamente el scale.data que estaba en sce.hash (como assays(sce.hash)[["scaledata"]]), dice que no lo tiene en cuenta. Igual despues no lo necesito porque uso el assay de RNA y no mas el de HTO
DefaultAssay(SeuRaw_noNeg) <- "RNA"
rm(sce.hash)
```

```{r include=T}
 gc()
```

## Analyzing cells using RNA assay

### Do intra- or -inter sample cells have higher ribosomal percentage?

```{r hashed.doublets_vln_perc.rb}
ggcells(as.SingleCellExperiment(SeuRaw_noNeg)) + geom_violin(aes(x = state, y =
                                                                   percent.rb))
```

```{r save_RawNoNeg_object, include = T}
#saveRDS(SeuRaw_noNeg, paste0("Outputs/", date, "_SeuRaw_noNeg_", numDS, "DataSet.rds"))
```

## Keep singlets and intra-sample doublets

```{r singlet_intra.sample_doublets}
SeuObj.singlet <-
  subset(SeuRaw_noNeg, subset = hash.ID != "Doublet") # (date = 230524)
rm(SeuRaw_noNeg)
```

When removing inter-samples doublets, the 2 dopaminergic Cre cells from before are lost
```{r}
which(SeuObj.singlet$CreCells != "FALSE")
```


Dim of the object is `r dim(SeuObj.singlet)`\`

```{r include = T}
gc()
```

## Analyze data without integration

In the "predicted" plot, blue point are predicted intra-sample doublets

```{r, include=T}
#SeuObj.singlet.merge <- norm_seurat(SeuObj.singlet)
SeuObj.singlet.merge <- FindVariableFeatures(SeuObj.singlet)
SeuObj.singlet.merge@reductions <- list()
SeuObj.singlet.merge <-
  UMAP_maker(SeuObj.singlet.merge, dims = 50, k.param = 10)
SeuObj.singlet.merge <-
  RunTSNE(SeuObj.singlet.merge,
          dims = 1:4,
          perplexity = 100)
SeuObj.singlet.merge <-
  FindNeighbors(
    SeuObj.singlet.merge,
    dims = 50,
    k.param = 10,
    compute.SNN = F
  )
SeuObj.singlet.merge <-
  FindClusters(
    SeuObj.singlet.merge,
    resolution = 0.05,
    n.start = 100,
    n.iter = 10
  )
```

```{r, include = T}
#Add meaning to HTO as metadata columns
equiv_table <-
  data.frame(
    hash.ID = unique(SeuObj.singlet.merge@meta.data$hash.ID),
    Treat = equiv_values_Treat,
    Age = equiv_values_Age,
    colors = colors_hash
  )
cols_hash <- equiv_table$colors
names(cols_hash) <- equiv_table$hash.ID
print("this is equiv table")
equiv_table
new_meta_data <-
  left_join(select(SeuObj.singlet.merge@meta.data,  "seurat_clusters", "hash.ID") ,
            equiv_table)
SeuObj.singlet.merge@meta.data$Treat <- new_meta_data$Treat
SeuObj.singlet.merge@meta.data$Age <- new_meta_data$Age
```

```{r DimplotAssay, out.width= '150%', out.height= '70%', fig.asp=0.45}
colors_n <- rainbow(length(unique(SeuObj.singlet.merge@meta.data$seurat_clusters)))
names(colors_n) <- unique(SeuObj.singlet.merge@meta.data$seurat_clusters)
UMAP.singletA <-
  DimPlot(SeuObj.singlet.merge,
          group.by = "hash.ID",
          cells = sample(Cells(SeuObj.singlet.merge)),
          cols = cols_hash) # Hacer que el orden de aparicion de celulas sea aleatorio
UMAP.singletB <-
  DimPlot(SeuObj.singlet.merge,
          group.by = "seurat_clusters",
          cols =  colors_n,
          cells = sample(Cells(SeuObj.singlet.merge)))
## Plot which are predicted doublets:
UMAP.singletC <-
  DimPlot(SeuObj.singlet.merge, group.by = "predicted", order = T)
(UMAP.singletA +  UMAP.singletB + UMAP.singletC) &
  theme(legend.position = "none")
tSNE.singletA <-
  DimPlot(
    SeuObj.singlet.merge,
    reduction = "tsne",
    group.by = "hash.ID",
    cells = sample(Cells(SeuObj.singlet.merge)),
    cols = cols_hash
  ) + # Hacer que el orden de aparicion de celulas sea aleatorio 
guides(color = guide_legend(nrow = 4, override.aes = list(size=3)))
tSNE.singletB <-
  DimPlot(
    SeuObj.singlet.merge,
    reduction = "tsne",
    group.by = "seurat_clusters",
    cols =  colors_n,
    cells = sample(Cells(SeuObj.singlet.merge))
  )
## Plot which are predicted doublets:
tSNE.singletC <-
  DimPlot(SeuObj.singlet.merge,
          reduction = "tsne",
          group.by = "predicted", order = T) +
  guides(color = guide_legend(nrow = 2, override.aes = list(size=3)))
(tSNE.singletA +  tSNE.singletB + tSNE.singletC) &
  theme(legend.position = "bottom", legend.key.size = unit(0.1, "cm"))

SeuObj.singlet.merge$CouldBeADouble <- ifelse(SeuObj.singlet.merge$seurat_clusters %in% cluster_map, T, F)

```

```{r ASW_batch_noInt, include=TRUE}
# ElbowPlot(SeuObj.singlet.merge, ndims = 50)
obj_dims = 30 # Dimensions to use
d <-
  stats::dist(Embeddings(SeuObj.singlet.merge, reduction = "pca")[, 1:obj_dims], method = "euclidean") # Create Matrix of Euclidean distances between all pair cells
# There are 2 type of reductions to use: "pca" and "umap". "pca" reduction has 50 dimensions while "umap" reduction has 2 dimensions. Since clustering is based on "pca" reductions, we will use "pca" reductions to get the silhoutte score. Despite everthing, UMAP is just a method to vizualize what is hoded on the 50 PCAs. But lets double check in the Seurat Paper and in my presentations if the Clustering use UMAP reduction
batches <-
  tibble(CellNames = dimnames(SeuObj.singlet.merge)[[2]],
         Batch = SeuObj.singlet.merge@meta.data$hash.ID) # Set the Batch correspondence to each cell
batches <- mutate(batches, Batch_int = as.numeric(as.factor(Batch)))
sil_batch <-
  cluster::silhouette(x = batches$Batch_int - 1, dist = d) # Calculate SW for batch effect for each cell
rm(batches, d)
ASW_batch <-
  mean(tibble::as_tibble(sil_batch[,])$sil_width) # Calculate the Average Silhoutte 
```

#### The ASW (Average Silhouette Width) batch score (integration quality parameter) for the non integrated data is `r round(ASW_batch, 4)`

```{r, include=TRUE}
# ElbowPlot(SeuObj.singlet.merge, ndims = 50)
SeuObj.singlet.merge_b <- subset(SeuObj.singlet.merge, subset = Treat == "Ctrl")
obj_dims = 30 # Dimensions to use
d <-
  stats::dist(Embeddings(SeuObj.singlet.merge_b, reduction = "pca")[, 1:obj_dims], method = "euclidean") # Create Matrix of Euclidean distances between all pair cells
# There are 2 type of reductions to use: "pca" and "umap". "pca" reduction has 50 dimensions while "umap" reduction has 2 dimensions. Since clustering is based on "pca" reductions, we will use "pca" reductions to get the silhoutte score. Despite everthing, UMAP is just a method to vizualize what is hoded on the 50 PCAs. But lets double check in the Seurat Paper and in my presentations if the Clustering use UMAP reduction
batches <-
  tibble(CellNames = dimnames(SeuObj.singlet.merge_b)[[2]],
         Batch = SeuObj.singlet.merge_b@meta.data$hash.ID) # Set the Batch correspondence to each cell
batches <- mutate(batches, Batch_int = as.numeric(as.factor(Batch)))
sil_batch <-
  cluster::silhouette(x = batches$Batch_int - 1, dist = d) # Calculate SW for batch effect for each cell
rm(batches, d)
ASW_batch <-
  mean(tibble::as_tibble(sil_batch[,])$sil_width) # Calculate the Average Silhoutte 
```

#### The ASW (Average Silhouette Width) batch score (integration quality parameter) for the non integrated data ONLY on Ctrl samples (different months as "replicates" of clusterization) is `r round(ASW_batch, 4)`

## Keep only singlets

```{r, include=T}
SeuObj.singlet.merge <-
  subset(SeuObj.singlet.merge, subset = state == "singlet") # (date = 230524)
SeuObj.singlet.merge <- FindVariableFeatures(SeuObj.singlet.merge)
SeuObj.singlet.merge <- UMAP_maker(SeuObj.singlet.merge, dims = 50, k.param = 10)
SeuObj.singlet.merge <- RunTSNE(SeuObj.singlet.merge, dims = 1:4, perplexity = 100)
SeuObj.singlet.merge <-
  FindNeighbors(SeuObj.singlet.merge,
                dims = 50,
                k.param = 10,
                compute.SNN = F)
SeuObj.singlet.merge <-
  FindClusters(SeuObj.singlet.merge,
               resolution = 0.05,
               n.start = 100,
               n.iter = 10)
```

when removing intra-sample doublets the other 2 Cre+ cells are removed
```{r}
which(SeuObj.singlet.merge$CreCells != "FALSE")
```


```{r, out.width= '150%', out.height= '70%', fig.asp=0.45}
UMAP.singletA <-
  DimPlot(SeuObj.singlet.merge,
          group.by = "hash.ID",
          cells = sample(Cells(SeuObj.singlet.merge)),
          cols = cols_hash) # Hacer que el orden de aparicion de celulas sea aleatorio
UMAP.singletB <-
  DimPlot(SeuObj.singlet.merge,
          group.by = "seurat_clusters",
          cells = sample(Cells(SeuObj.singlet.merge)),
          cols = colors_n)
## Plot which are predicted doublets:
UMAP.singletC <-
  DimPlot(SeuObj.singlet.merge, group.by = "predicted", order = T)
(UMAP.singletA +  UMAP.singletB + UMAP.singletC) &
  theme(legend.position = "none")
tSNE.singletA <-
  DimPlot(
    SeuObj.singlet.merge,
    reduction = "tsne",
    group.by = "hash.ID",
    cells = sample(Cells(SeuObj.singlet.merge)),
    cols = cols_hash
  ) + # Hacer que el orden de aparicion de celulas sea aleatorio
  guides(color = guide_legend(nrow = 4, override.aes = list(size=3)))
tSNE.singletB <-
  DimPlot(
    SeuObj.singlet.merge,
    reduction = "tsne",
    group.by = "seurat_clusters",
    cells = sample(Cells(SeuObj.singlet.merge)),
    cols = colors_n
  )
## Plot which are predicted doublets:
tSNE.singletC <-
  DimPlot(SeuObj.singlet.merge,
          reduction = "tsne",
          group.by = "predicted", order = T) +
  guides(color = guide_legend(nrow = 2, override.aes = list(size=3)))
(tSNE.singletA +  tSNE.singletB + tSNE.singletC) &
  theme(legend.position = "bottom", legend.key.size = unit(0.1, "cm"))
```

```{r}
Table_no_filter <- table(SeuObj.singlet.merge$seurat_clusters, SeuObj.singlet.merge$hash.ID)
Table_no_filter
```

```{r, include=TRUE}
# ElbowPlot(SeuObj.singlet.merge, ndims = 50)
obj_dims = 30 # Dimensions to use
d <-
  stats::dist(Embeddings(SeuObj.singlet.merge, reduction = "pca")[, 1:obj_dims], method = "euclidean") # Create Matrix of Euclidean distances between all pair cells
# There are 2 type of reductions to use: "pca" and "umap". "pca" reduction has 50 dimensions while "umap" reduction has 2 dimensions. Since clustering is based on "pca" reductions, we will use "pca" reductions to get the silhoutte score. Despite everthing, UMAP is just a method to vizualize what is hoded on the 50 PCAs. But lets double check in the Seurat Paper and in my presentations if the Clustering use UMAP reduction
batches <-
  tibble(CellNames = dimnames(SeuObj.singlet.merge)[[2]],
         Batch = SeuObj.singlet.merge@meta.data$hash.ID) # Set the Batch correspondence to each cell
batches <- mutate(batches, Batch_int = as.numeric(as.factor(Batch)))
sil_batch <-
  cluster::silhouette(x = batches$Batch_int - 1, dist = d) # Calculate SW for batch effect for each cell
rm(batches, d)
ASW_batch <-
  mean(tibble::as_tibble(sil_batch[,])$sil_width) # Calculate the Average Silhoutte 
```

#### The ASW (Average Silhouette Width) batch score (integration quality parameter) for the non integrated data is `r round(ASW_batch, 4)`

```{r, include=TRUE}
# ElbowPlot(SeuObj.singlet.merge, ndims = 50)
SeuObj.singlet.merge_b <- subset(SeuObj.singlet.merge, subset = Treat == "Ctrl")
obj_dims = 30 # Dimensions to use
d <-
  stats::dist(Embeddings(SeuObj.singlet.merge_b, reduction = "pca")[, 1:obj_dims], method = "euclidean") # Create Matrix of Euclidean distances between all pair cells
# There are 2 type of reductions to use: "pca" and "umap". "pca" reduction has 50 dimensions while "umap" reduction has 2 dimensions. Since clustering is based on "pca" reductions, we will use "pca" reductions to get the silhoutte score. Despite everthing, UMAP is just a method to vizualize what is hoded on the 50 PCAs. But lets double check in the Seurat Paper and in my presentations if the Clustering use UMAP reduction
batches <-
  tibble(CellNames = dimnames(SeuObj.singlet.merge_b)[[2]],
         Batch = SeuObj.singlet.merge_b@meta.data$hash.ID) # Set the Batch correspondence to each cell
batches <- mutate(batches, Batch_int = as.numeric(as.factor(Batch)))
sil_batch <-
  cluster::silhouette(x = batches$Batch_int - 1, dist = d) # Calculate SW for batch effect for each cell
rm(batches, d)
ASW_batch <-
  mean(tibble::as_tibble(sil_batch[,])$sil_width) # Calculate the Average Silhoutte 
```

#### The ASW (Average Silhouette Width) batch score (integration quality parameter) for the non integrated data ONLY on Ctrl samples (different months as "replicates" of clusterization) is `r round(ASW_batch, 4)`

```{r DataSummary1, include=T}
# Data summary
# print("data summary singlet1")
# data_summary <-
#   data.frame(
#     "value" = c(
#       "date" = date,
#       "path" = paste0(wd, FilePath, FileName),
#       "object_name" = "SeuObj.singlet.merge",
#       "filtered" = "no",
#       "nr_cells" = dim(SeuObj.singlet.merge)[2],
#       "demultiplexed" = "yes",
#       "demux_positive.quantile" = 0.99,
#       "norm.meth" =  "CLR",
#       # Change this if needed
#       HTO_classification,
#       "integrated" = "no",
#       "ASW_batch" = round(ASW_batch, 4) ,
#       "ASW_dims" = obj_dims,
#       "nr_clusters" = length(levels(
#         SeuObj.singlet.merge@meta.data$seurat_clusters
#       ))
#     )
#   ) %>% mutate("vars" = rownames(.), .before = value)  %>% `rownames<-`(NULL)
# neigh_params <-
#   SeuObj.singlet.merge$FindNeighbors.RNA.pca@params %>% unlist() %>% as.data.frame() %>% `colnames<-`("value")  %>% mutate("vars" = rownames(.), .before = value)  %>% `rownames<-`(NULL) %>% rbind(data.frame("vars" =  "neigh.params", "value" = NA), .)
# clust_params <-
#   SeuObj.singlet.merge$FindClusters@params %>% unlist() %>% as.data.frame() %>% `colnames<-`("value")  %>% mutate("vars" = rownames(.), .before = value)  %>% `rownames<-`(NULL) %>% rbind(data.frame("vars" =  "clust.params", "value" = NA), .)
# norm_params <-
#   SeuObj.singlet.merge$NormalizeData.RNA@params %>% unlist() %>% as.data.frame() %>% `colnames<-`("value")  %>% mutate("vars" = rownames(.), .before = value)  %>% `rownames<-`(NULL) %>% rbind(data.frame("vars" =  "norm.params", "value" = NA), .)
# variabFeat_params <-
#   SeuObj.singlet.merge$FindVariableFeatures.RNA@params %>% unlist() %>% .[-c(5, 6)] %>% as.data.frame() %>% t() %>% as.data.frame() %>% `colnames<-`("value")  %>% mutate("vars" = rownames(.), .before = value)  %>% `rownames<-`(NULL) %>% rbind(data.frame("vars" =  "varFeat.params", "value" = NA), .)
# data_summary <-
#   rbind(data_summary, neigh_params) %>% rbind(clust_params) %>% rbind(norm_params) %>% rbind(variabFeat_params)
# data_summary
```

### Color the UMAPs by library size:

The plots differ in how the scale is adjusted.

```{r,  out.width= '100%', out.height= '75%', fig.asp=1.5}
thfp <-
  theme(plot.title = element_text(size = 9), legend.text = element_text(size = 4), axis.text = element_text(size = 4), axis.title = element_text(size = 4))
fp1 <-
  FeaturePlot(object = SeuObj.singlet.merge,
              features = "UMIs.p.cell",
              cells =  sample(Cells(SeuObj.singlet.merge))) + scale_colour_gradient(low = "lightblue", high = "darkblue") + 
  labs(title = "All; no lim on lbsz")+ thfp

fp2 <-
  FeaturePlot(object = SeuObj.singlet.merge,
              features = "UMIs.p.cell",
              cells =  sample(Cells(SeuObj.singlet.merge))) + scale_colour_gradient(
                limits = c(0, 10000),
                low = "lightblue",
                high = "darkblue"
              ) +  labs(title = "All; scale lim to 10000") + thfp

fp3 <-
  FeaturePlot(object = SeuObj.singlet.merge,
              features = "UMIs.p.cell",
              cells =  sample(Cells(SeuObj.singlet.merge))) + scale_colour_gradient(
                limits = c(0, 20000),
                low = "lightblue",
                high = "darkblue"
              ) + labs(title = "All; scale lim to 20000") +thfp

fp4 <-
  FeaturePlot(
    object = subset(SeuObj.singlet.merge, subset = Treat == "Ctrl"),
    features = "UMIs.p.cell",
    cells =  sample(Cells(subset(
      SeuObj.singlet.merge, subset = Treat == "Ctrl"
    )))
  ) + scale_colour_gradient(
    limits = c(0, 20000),
    low = "lightblue",
    high = "darkblue"
  ) + labs(title = "Ctrl; scale lim to 20000") + thfp

fp5 <-
  FeaturePlot(
    object = subset(SeuObj.singlet.merge, subset = Treat == "PD"),
    features = "UMIs.p.cell",
    cells =  sample(Cells(subset(
      SeuObj.singlet.merge, subset = Treat == "PD"
    )))
  ) + scale_colour_gradient(
    limits = c(0, 20000),
    low = "lightblue",
    high = "darkblue"
  ) + labs(title = "PD; scale lim to 20000") + thfp

fp6 <-
  FeaturePlot(
    object = subset(SeuObj.singlet.merge, subset = Treat == "Ctrl" &
                      Age == "1mo"),
    features = "UMIs.p.cell",
    cells =  sample(Cells(
      subset(SeuObj.singlet.merge, subset = Treat == "Ctrl" &
               Age == "1mo")
    ))
  ) + scale_colour_gradient(
    limits = c(0, 20000),
    low = "lightblue",
    high = "darkblue"
  ) + labs(title = "Ctrl 1mo; scale lim to 20000") + thfp

fp7 <-
  FeaturePlot(
    object = subset(SeuObj.singlet.merge, subset = Treat == "Ctrl" &
                      Age == "2mo"),
    features = "UMIs.p.cell",
    cells =  sample(Cells(
      subset(SeuObj.singlet.merge, subset = Treat == "Ctrl" &
               Age == "2mo")
    ))
  ) + scale_colour_gradient(
    limits = c(0, 20000),
    low = "lightblue",
    high = "darkblue"
  ) + labs(title = "Ctrl 2mo; scale lim to 20000") + thfp

plot_grid(fp1, fp2, fp3, fp4, fp5, fp6, fp7, NULL, ncol = 2)
```

### Filter-out cells having less that 1000 and more than 60000 UMI counts

I decide not to filter cell for their ribosomal protein content. Here
are some paper in which the ribosomal proteins were differentially
expressed in PD. In some papers these genes are up-regulated and in
other downregulated, depending the case:

-   @garcia-esparcia2015 found 12 Rpl/Rps genes downregulated in SN in
    stages 3-4

-   @hemmings2022 Found a module of Rpl/Rps genes associated with a PD
    and PTSD population. Also they cite a paper in which Rpl6 was
    associated with early onset of PD

-   @jiao2023 have a nice review of ribosome biogenesis

-   @zhong2020 found that Rpl3, Rpl6, Rps27a, Rps29 were up-regulated in
    a snRNA-seq experiment in the midbrain of a PD mouse

However, cell with higher ribosomal content have also the lowest number
of active genes. This could be either some sort of bias given by the
manipulation of cells or maybe those are cells that have some sort of
metabolic disregulation becaus eof the PD treatment.

```{r, out.width= '100%', out.height= '100%'}
UMIs.p.cell_lowthreshold = 1000 #2000
UMIs.p.cell_highthreshold = 60000
percent.rb_threshold = Inf
SeuObj.singlet.merge <-
  subset(
    SeuObj.singlet.merge,
    subset = UMIs.p.cell > UMIs.p.cell_lowthreshold &
      UMIs.p.cell < UMIs.p.cell_highthreshold &
      percent.rb < percent.rb_threshold
  )

## Find clusters and color UMAP by clusters with Louvain and a high number of iterations ####
UMAP.SeuInt1 <-
  DimPlot(SeuObj.singlet.merge, group.by = "hash.ID", cells = sample(Cells(SeuObj.singlet.merge)), cols = cols_hash) + guides(color = guide_legend(override.aes = list(size = 1), ncol = 2))
UMAP.SeuInt2 <-
  DimPlot(SeuObj.singlet.merge, group.by = "seurat_clusters", cols =  colors_n, cells = sample(Cells(SeuObj.singlet.merge)))
(UMAP.SeuInt1 +  UMAP.SeuInt2) &
  theme(legend.position = "bottom",
        legend.key.size = unit(0.1, "cm"))

## Plot QC graphics per cluster ####
QC_plots(SeuObj.singlet.merge, "", colors_n)
```
```{r}
Table_filter <- table(SeuObj.singlet.merge$seurat_clusters, SeuObj.singlet.merge$hash.ID)
Table_filter
```
Difference
```{r}
Table_filter - Table_no_filter
```

```{r}
SeuObj.singlet.merge <- FindVariableFeatures(SeuObj.singlet.merge)
SeuObj.singlet.merge <- UMAP_maker(SeuObj.singlet.merge, dims = 50, k.param = 10)
SeuObj.singlet.merge <- RunTSNE(SeuObj.singlet.merge, dims = 1:4, perplexity = 100)
SeuObj.singlet.merge <-
  FindNeighbors(SeuObj.singlet.merge,
                dims = 50,
                k.param = 10,
                compute.SNN = F)
SeuObj.singlet.merge <-
  FindClusters(SeuObj.singlet.merge,
               resolution = 0.05,
               n.start = 100,
               n.iter = 10)
```


```{r}
## Find clusters and color UMAP by clusters with Louvain and a high number of iterations ####
colors_n <- rainbow(length(unique(SeuObj.singlet.merge@meta.data$seurat_clusters)))
names(colors_n) <- unique(SeuObj.singlet.merge@meta.data$seurat_clusters)

UMAP.SeuInt1_b <-
  DimPlot(SeuObj.singlet.merge, group.by = "hash.ID", cells = sample(Cells(SeuObj.singlet.merge)), cols = cols_hash) + guides(color = guide_legend(override.aes = list(size = 1), ncol = 2))
UMAP.SeuInt2_b <-
  DimPlot(SeuObj.singlet.merge, group.by = "seurat_clusters", cols =  colors_n, cells = sample(Cells(SeuObj.singlet.merge)))
(UMAP.SeuInt1_b +  UMAP.SeuInt2_b + UMAP.SeuInt2) &
  theme(legend.position = "bottom",
        legend.key.size = unit(0.1, "cm"))

## Plot QC graphics per cluster ####
QC_plots(SeuObj.singlet.merge, "", colors_n)
```


```{r, include=TRUE}
# ElbowPlot(SeuObj.singlet.merge, ndims = 50)
obj_dims = 30 # Dimensions to use
d <-
  stats::dist(Embeddings(SeuObj.singlet.merge, reduction = "pca")[, 1:obj_dims], method = "euclidean") # Create Matrix of Euclidean distances between all pair cells
# There are 2 type of reductions to use: "pca" and "umap". "pca" reduction has 50 dimensions while "umap" reduction has 2 dimensions. Since clustering is based on "pca" reductions, we will use "pca" reductions to get the silhoutte score. Despite everthing, UMAP is just a method to vizualize what is hoded on the 50 PCAs. But lets double check in the Seurat Paper and in my presentations if the Clustering use UMAP reduction
batches <-
  tibble(CellNames = dimnames(SeuObj.singlet.merge)[[2]],
         Batch = SeuObj.singlet.merge@meta.data$hash.ID) # Set the Batch correspondence to each cell
batches <- mutate(batches, Batch_int = as.numeric(as.factor(Batch)))
sil_batch <-
  cluster::silhouette(x = batches$Batch_int - 1, dist = d) # Calculate SW for batch effect for each cell
rm(batches, d)
ASW_batch <-
  mean(tibble::as_tibble(sil_batch[,])$sil_width) # Calculate the Average Silhoutte 
```

#### The ASW (Average Silhouette Width) batch score (integration quality parameter) for the non integrated data is `r round(ASW_batch, 4)`

```{r, include=TRUE}
# ElbowPlot(SeuObj.singlet.merge, ndims = 50)
SeuObj.singlet.merge_b <- subset(SeuObj.singlet.merge, subset = Treat == "Ctrl")
obj_dims = 30 # Dimensions to use
d <-
  stats::dist(Embeddings(SeuObj.singlet.merge_b, reduction = "pca")[, 1:obj_dims], method = "euclidean") # Create Matrix of Euclidean distances between all pair cells
# There are 2 type of reductions to use: "pca" and "umap". "pca" reduction has 50 dimensions while "umap" reduction has 2 dimensions. Since clustering is based on "pca" reductions, we will use "pca" reductions to get the silhoutte score. Despite everthing, UMAP is just a method to vizualize what is hoded on the 50 PCAs. But lets double check in the Seurat Paper and in my presentations if the Clustering use UMAP reduction
batches <-
  tibble(CellNames = dimnames(SeuObj.singlet.merge_b)[[2]],
         Batch = SeuObj.singlet.merge_b@meta.data$hash.ID) # Set the Batch correspondence to each cell
batches <- mutate(batches, Batch_int = as.numeric(as.factor(Batch)))
sil_batch <-
  cluster::silhouette(x = batches$Batch_int - 1, dist = d) # Calculate SW for batch effect for each cell
rm(batches, d)
ASW_batch <-
  mean(tibble::as_tibble(sil_batch[,])$sil_width) # Calculate the Average Silhoutte 
```

#### The ASW (Average Silhouette Width) batch score (integration quality parameter) for the non integrated data ONLY on Ctrl samples (different months as "replicates" of clusterization) is `r round(ASW_batch, 4)`

```{r}
Table_filter <- table(SeuObj.singlet.merge$seurat_clusters, SeuObj.singlet.merge$hash.ID)
Table_filter
```



```{r FilterCell_QC, out.width= '100%', out.height= '100%'}
# DefaultAssay(SeuObj.singlet.merge) <- "RNA"
# SeuInt <- FindVariableFeatures(SeuObj.singlet.merge)
# SeuInt <- UMAP_maker(SeuObj.singlet.merge, 50, 10)
# 
# ## Find clusters and color UMAP by clusters with Louvain and a high number of iterations ####
# UMAP.SeuInt1 <-
#   DimPlot(SeuInt, group.by = "hash.ID", cells = sample(Cells(SeuInt)), cols = cols_hash) + guides(color = guide_legend(override.aes = list(size = 1), ncol = 2))
# UMAP.SeuInt2 <-
#   DimPlot(SeuInt, group.by = "seurat_clusters", cols =  colors_n, cells = sample(Cells(SeuInt)))
# (UMAP.SeuInt1 +  UMAP.SeuInt2) &
#   theme(legend.position = "bottom",
#         legend.key.size = unit(0.1, "cm"))
# 
# ## Plot QC graphics per cluster ####
# QC_plots(SeuInt, "", colors_n)
```

## Integrate Samples

The way to do the integration is the following:

-   Select the features to integrate: this are the joint 2000 top
    variable features in all samples. The variable features are already
    computed separately on each sample. Here, features are positively
    ranked for being variable in all samples at the same time.

-   For each pair of samples, find the pair of cells between them that
    can match the best in the reduced PCA space (anchors) of each
    sample. It is like finding the *k* nearest neighborhood of a cell in
    the PCA projection of the paired sample

-   Do a hierarchical clustering between samples (using a specific
    distance calculation based on anchors) to get a guide tree to make
    the integration of a pair of samples. For each pair: Integrate both
    samples subtracting the expression values of the selected features
    from one member of the anchor to the other member. Without changing
    the values of the cells that couldn't be anchored? (mirar
    anotaciones). A weight is also applied to penalize the distance
    between each anchor. Repeat these steps using the newly generated
    matrix and the following sample in the guide tree to integrate.

    In my particular case, I integrated using the selected variable
    features plus some neuronal, glial, and capilar cell-markers of
    interest, plus the transgenes.

-   Values for k.anchor, k.filter, k.score, max.features and k.weight
    are obtained by the ones tested in the @otero-garcia2022 paper DS
    (from Inma Cobos lab) which retrieved best ASW batch scores (see
    Alzheimer_project/Scripts/scRNAseq_integration_RCPA_labmeet.html to
    see the values)

```{r Integrate, include = T}
print("start integration")
# Prepare the objects to integrate
#SeuObj.singlet <- subset(SeuObj.singlet, subset = state == "singlet")
SeuObj.singlet.split <-
  SplitObject(SeuObj.singlet.merge, split.by = "hash.ID") # Create a list of Seurat Objects to integrate
SeuObj.singlet.split <-
  lapply(SeuObj.singlet.split, norm_seurat) # Normalize and select variable features
SeuIntFeatures <-
  SelectIntegrationFeatures(SeuObj.singlet.split, nfeatures = 2000) # Features (genes) to use in the integration
SeuObj.singlet.split <-
  lapply(SeuObj.singlet.split, preprocess_RPCA, SeuIntFeatures) # Pre-process data: Scale data and run PCA

SNmark <-
  c("Th",
    "Slc6a3",
    "Pitx3",
    "Smad3",
    "Neurod6",
    "Slc18a3",
    "Slc18a2",
    "Ddc") # Removed markers:  "Adra1a" "Ovgp1", "Smad9", "Gal", "Galr1", "Npy", falta "Galr3" y "Galr2"
Neuron <- c("Rbfox3") # General neurons
Ex <- c("Slc17a6", "Slc17a7") # Excitatory neurons
Inh <- c("Gad1") # Inhibitory neuron
Astrocytes <- c("Gfap", "Gja1") # Removed markers: Blbp = Fabp7;
Microglia <- c("C1qb") # Removed markers: Aif1 = Iba1
OPC <- c("Olig1", "Tnr")
Oligodendrocyte <- c("Tfrc", "Plp1", "Mbp") # Tfrc = Tfr
Fibro <- c("Dcn") # Fibroblast
Capil <- c("Flt1", "Rgs5")
markers <- c("EGFP-KASH", "hSNCA-A53T", "mCherry",  "Snca")
# GliaMark<-c("Aldh1l1", "Aldoc", "Olig2", "Cnp")

# Cell-cell pairings (anchors) to do the integration
# Method: Reciprocal PCA
anchors <-
  FindIntegrationAnchors(
    SeuObj.singlet.split,
    reduction = "rpca",
    dims = 1:obj_dims,
    #anchor.features = SeuIntFeatures,
    anchor.features =  c(
      SeuIntFeatures,
      SNmark,
      Neuron,
      Ex,
      Inh,
      Fibro,
      Astrocytes,
      Microglia,
      Oligodendrocyte,
      OPC,
      Capil,
      markers
    ),
    scale = F,
    k.anchor = 10,
    k.filter = 100,
    k.score = 20,
    max.features = 100
  ) # FindAnchors
# Features to include in the integration (markers of different cell types, including Substancia Nigra (SN), extracted from PanglaoDb)

# Integrate data including in the final matrix some marker genes
SeuInt <-
  IntegrateData(
    anchorset = anchors,
    dims = 1:obj_dims,
    k.weight = 20,
    features.to.integrate = c(
      SeuIntFeatures,
      SNmark,
      Neuron,
      Ex,
      Inh,
      Fibro,
      Astrocytes,
      Microglia,
      Oligodendrocyte,
      OPC,
      Capil,
      markers
    )
  )
rm(anchors, SeuObj.singlet.split)
```

```{r include = T}
gc()
```

```{r HTOmeaning, include = T}
# #Add meaning to HTO as metadata columns
# equiv_table <-
#   data.frame(
#     hash.ID = unique(SeuInt@meta.data$hash.ID),
#     Treat = equiv_values_Treat,
#     Age = equiv_values_Age
#   )
# print("this is equiv table")
# equiv_table
# new_meta_data <-
#   left_join(select(SeuInt@meta.data, "hash.ID") ,
#             equiv_table)
# SeuInt@meta.data$Treat <- new_meta_data$Treat
# SeuInt@meta.data$Age <- new_meta_data$Age
```

```{r ASW_batch_Int, include=TRUE}
SeuInt <- UMAP_maker(SeuInt, dims = 50, k.param = 10)
SeuInt <- RunTSNE(SeuInt, dims = 1:4, perplexity = 100)

# Obtain the ASW (Average Silhouette Width) batch score (integration quality parameter)
# ElbowPlot(SeuInt, ndims = 50)
obj_dims = 30 # Dimensions to use
d <-
  stats::dist(Embeddings(SeuInt, reduction = "pca")[, 1:obj_dims], method = "euclidean") # Create Matrix of Euclidean distances between all pair cells
batches <-
  tibble(CellNames = dimnames(SeuInt)[[2]],
         Batch = SeuInt@meta.data$hash.ID) # Set the Batch correspondence to each cell
batches <- mutate(batches, Batch_int = as.numeric(as.factor(Batch)))
sil_batch <-
  cluster::silhouette(x = batches$Batch_int - 1, dist = d) # Calculate SW for batch effect for each cell
rm(batches, d)
ASW_batch <-
  mean(tibble::as_tibble(sil_batch[,])$sil_width) # Calculate the Average Silhoutte 
```

#### The ASW (Average Silhouette Width) batch score (integration quality parameter) for the integrated data is `r round(ASW_batch, 4)`

```{r, include=TRUE}
# ElbowPlot(SeuInt, ndims = 50)
SeuInt_b <- subset(SeuInt, subset = Treat == "Ctrl")
obj_dims = 30 # Dimensions to use
d <-
  stats::dist(Embeddings(SeuInt_b, reduction = "pca")[, 1:obj_dims], method = "euclidean") # Create Matrix of Euclidean distances between all pair cells
# There are 2 type of reductions to use: "pca" and "umap". "pca" reduction has 50 dimensions while "umap" reduction has 2 dimensions. Since clustering is based on "pca" reductions, we will use "pca" reductions to get the silhoutte score. Despite everthing, UMAP is just a method to vizualize what is hoded on the 50 PCAs. But lets double check in the Seurat Paper and in my presentations if the Clustering use UMAP reduction
batches <-
  tibble(CellNames = dimnames(SeuInt_b)[[2]],
         Batch = SeuInt_b@meta.data$hash.ID) # Set the Batch correspondence to each cell
batches <- mutate(batches, Batch_int = as.numeric(as.factor(Batch)))
sil_batch <-
  cluster::silhouette(x = batches$Batch_int - 1, dist = d) # Calculate SW for batch effect for each cell
rm(batches, d)
ASW_batch <-
  mean(tibble::as_tibble(sil_batch[,])$sil_width) # Calculate the Average Silhoutte 
```

#### The ASW (Average Silhouette Width) batch score (integration quality parameter) for the non integrated data ONLY on Ctrl samples (different months as "replicates" of clusterization) is `r round(ASW_batch, 4)`

### Find clusters with Louvain's algorithm

See also which clusters have a higher number of intra-sample doublets
and analyze some quality control measures of each of the clusters.

```{r, include=T}
SeuInt <-
  FindNeighbors(SeuInt,
                dims = 50,
                k.param = 10,
                compute.SNN = F)
SeuInt <-
  FindClusters(SeuInt,
               resolution = 0.05,
               #graph.name = "integrated_snn",
               #group.singletons = F, NO HAY SINGLETONS
               n.start = 100,
               n.iter = 10)

SeuInt <- RunUMAP(SeuInt, dims = 1:50)

#SeuInt <- RunUMAP(SeuInt, graph = "integrated_snn")
#(SeuInt, dims = 50, k.param = 10, graph = "integrated_snn")

```

```{r,  out.width= '150%', out.height= '70%', fig.asp=0.5}
colors_n <- rainbow(length(unique(SeuInt@meta.data$seurat_clusters)))
names(colors_n) <- unique(SeuInt@meta.data$seurat_clusters)
UMAP.SeuInt1 <-
  DimPlot(SeuInt, group.by = "hash.ID", cells =  sample(Cells(SeuInt)), cols = cols_hash)
UMAP.SeuInt2 <-
  DimPlot(SeuInt, group.by = "seurat_clusters", cols =  colors_n, cells =  sample(Cells(SeuInt)), label = T)
(UMAP.SeuInt1 +  UMAP.SeuInt2) &
  theme(legend.position = "none")
tSNE.SeuInt1 <-
  DimPlot(
    SeuInt,
    reduction = "tsne",
    group.by = "hash.ID",
    cells =  sample(Cells(SeuInt)),
    cols = cols_hash
  ) + guides(color = guide_legend(nrow = 4, override.aes = list(size=3)))
tSNE.SeuInt2 <-
  DimPlot(
    SeuInt,
    reduction = "tsne",
    group.by = "seurat_clusters",
    cols = colors_n,
    cells =  sample(Cells(SeuInt))
  )
(tSNE.SeuInt1 +  tSNE.SeuInt2) &
  theme(legend.position = "bottom", legend.key.size = unit(0.1, "cm"))
```

```{r, out.width= '100%', out.height= '100%'}
# Plot QC graphics per cluster ####
QC_plots(SeuInt, "", colors_n)
```

#### How many cells for each cluster?

```{r}
table(SeuInt$seurat_clusters, SeuInt$Treat, SeuInt$Age)
```

Delate cluster 18 and 19:

```{r}
SeuInt2 <- subset(SeuInt, subset = seurat_clusters != "18" & seurat_clusters != "19")
table(SeuInt2$seurat_clusters, SeuInt2$Treat, SeuInt2$Age)
SeuInt2$seurat_clusters <- factor(SeuInt2$seurat_clusters, levels = unique(SeuInt2$seurat_clusters))
Idents(SeuInt2) %>% unique()
```


## Re clusterize

```{r, include=T}
SeuInt2 <- UMAP_maker(SeuInt2, dims = 50, k.param = 10)
```

```{r,  out.width= '150%', out.height= '70%', fig.asp=0.5}
colors_n <- rainbow(length(unique(SeuInt2@meta.data$seurat_clusters)))
names(colors_n) <- unique(SeuInt2@meta.data$seurat_clusters)
UMAP.SeuInt21 <-
  DimPlot(SeuInt2, group.by = "hash.ID", cells =  sample(Cells(SeuInt2)), cols = cols_hash)
UMAP.SeuInt22 <-
  DimPlot(SeuInt2, group.by = "seurat_clusters", cols =  colors_n, cells =  sample(Cells(SeuInt2)), label = T)
(UMAP.SeuInt21 +  UMAP.SeuInt2 + UMAP.SeuInt22) &
  theme(legend.position = "none")
tSNE.SeuInt21 <-
  DimPlot(
    SeuInt2,
    reduction = "tsne",
    group.by = "hash.ID",
    cells =  sample(Cells(SeuInt2)),
    cols = cols_hash
  ) + guides(color = guide_legend(nrow = 4, override.aes = list(size=3)))
tSNE.SeuInt22 <-
  DimPlot(
    SeuInt2,
    reduction = "tsne",
    group.by = "seurat_clusters",
    cols = colors_n,
    cells =  sample(Cells(SeuInt2))
  )
(tSNE.SeuInt21 +  tSNE.SeuInt22) &
  theme(legend.position = "bottom", legend.key.size = unit(0.1, "cm"))
```

Re-clusterize

```{r}
SeuInt2 <-
  FindClusters(SeuInt2,
               resolution = 0.1,
               #graph.name = "integrated_snn",
               #group.singletons = F, NO HAY SINGLETONS
               n.start = 100,
               n.iter = 10)
```


```{r UMAP-TSNE-QC_Int,  out.width= '150%', out.height= '70%', fig.asp=0.5}
colors_n <- rainbow(length(unique(SeuInt2@meta.data$seurat_clusters)))
names(colors_n) <- unique(SeuInt2@meta.data$seurat_clusters)
UMAP.SeuInt21b <-
  DimPlot(SeuInt2, group.by = "hash.ID", cells =  sample(Cells(SeuInt2)), cols = cols_hash)
UMAP.SeuInt22b <-
  DimPlot(SeuInt2, group.by = "seurat_clusters", cols =  colors_n, cells =  sample(Cells(SeuInt2)), label = T)
(UMAP.SeuInt21b +  UMAP.SeuInt22 + UMAP.SeuInt22b) &
  theme(legend.position = "none")
tSNE.SeuInt21 <-
  DimPlot(
    SeuInt2,
    reduction = "tsne",
    group.by = "hash.ID",
    cells =  sample(Cells(SeuInt2)),
    cols = cols_hash
  ) + guides(color = guide_legend(nrow = 4, override.aes = list(size=3)))
tSNE.SeuInt22 <-
  DimPlot(
    SeuInt2,
    reduction = "tsne",
    group.by = "seurat_clusters",
    cols = colors_n,
    cells =  sample(Cells(SeuInt2))
  )
## Plot which are predicted doublets:
(tSNE.SeuInt21 +  tSNE.SeuInt22) &
  theme(legend.position = "bottom", legend.key.size = unit(0.1, "cm"))
```


```{r}
table(SeuInt2$seurat_clusters, SeuInt2$Treat, SeuInt2$Age)
```

Delate cluster 20:

```{r}
SeuInt3 <- subset(SeuInt2, subset = seurat_clusters != "20")
table(SeuInt3$seurat_clusters, SeuInt3$Treat, SeuInt3$Age)
SeuInt3$seurat_clusters <- factor(SeuInt3$seurat_clusters, levels = unique(SeuInt3$seurat_clusters))
Idents(SeuInt3) %>% unique()
```


## Re clusterize

```{r, include=T}
# SeuInt3 <- FindVariableFeatures(SeuInt3)
# SeuInt3 <- UMAP_maker(SeuInt2, dims = 50, k.param = 10)
```

```{r,  out.width= '150%', out.height= '70%', fig.asp=0.5}
colors_n <- rainbow(length(unique(SeuInt3@meta.data$seurat_clusters)))
names(colors_n) <- unique(SeuInt3@meta.data$seurat_clusters)
UMAP.SeuInt31 <-
  DimPlot(SeuInt3, group.by = "hash.ID", cells =  sample(Cells(SeuInt3)), cols = cols_hash)
UMAP.SeuInt32 <-
  DimPlot(SeuInt3, group.by = "seurat_clusters", cols =  colors_n, cells =  sample(Cells(SeuInt3)), label = T)
(UMAP.SeuInt31 +  UMAP.SeuInt22b + UMAP.SeuInt32) &
  theme(legend.position = "none")
tSNE.SeuInt31 <-
  DimPlot(
    SeuInt3,
    reduction = "tsne",
    group.by = "hash.ID",
    cells =  sample(Cells(SeuInt3)),
    cols = cols_hash
  ) + guides(color = guide_legend(nrow = 4, override.aes = list(size=3)))
tSNE.SeuInt32 <-
  DimPlot(
    SeuInt3,
    reduction = "tsne",
    group.by = "seurat_clusters",
    cols = colors_n,
    cells =  sample(Cells(SeuInt3))
  )
(tSNE.SeuInt31 +  tSNE.SeuInt32) &
  theme(legend.position = "bottom", legend.key.size = unit(0.1, "cm"))
```


Join cluster 19 with cluster 1:
```{r}
SeuInt3$seurat_clusters <- gsub("19", "1", SeuInt3$seurat_clusters)
SeuInt3$seurat_clusters <- factor(SeuInt3$seurat_clusters, levels = unique(SeuInt3$seurat_clusters))
Idents(SeuInt3) <- SeuInt3$seurat_clusters
```

```{r,  out.width= '150%', out.height= '70%', fig.asp=0.5}
colors_n <- rainbow(length(unique(SeuInt3@meta.data$seurat_clusters)))
names(colors_n) <- unique(SeuInt3@meta.data$seurat_clusters)
UMAP.SeuInt31b <-
  DimPlot(SeuInt3, group.by = "hash.ID", cells =  sample(Cells(SeuInt3)), cols = cols_hash)
UMAP.SeuInt32b <-
  DimPlot(SeuInt3, group.by = "seurat_clusters", cols =  colors_n, cells =  sample(Cells(SeuInt3)), label = T)
(UMAP.SeuInt31b +  UMAP.SeuInt32 + UMAP.SeuInt32b) &
  theme(legend.position = "none")
tSNE.SeuInt31 <-
  DimPlot(
    SeuInt3,
    reduction = "tsne",
    group.by = "hash.ID",
    cells =  sample(Cells(SeuInt3)),
    cols = cols_hash
  ) + guides(color = guide_legend(nrow = 4, override.aes = list(size=3)))
tSNE.SeuInt32 <-
  DimPlot(
    SeuInt3,
    reduction = "tsne",
    group.by = "seurat_clusters",
    cols = colors_n,
    cells =  sample(Cells(SeuInt3))
  )
## Plot which are predicted doublets:
(tSNE.SeuInt31 +  tSNE.SeuInt32) &
  theme(legend.position = "bottom", legend.key.size = unit(0.1, "cm"))
```

```{r, out.width= '100%', out.height= '100%'}
# Plot QC graphics per cluster ####
QC_plots(SeuInt3, "", colors_n)
```



### Re-do plot:

```{r UMAP-TSNE-QC_Int_singlets, out.width= '100%', out.height= '100%'}
UMAP.SeuInt31 <-
  DimPlot(SeuInt3, group.by = "hash.ID", cells =  sample(Cells(SeuInt3)), cols = cols_hash) + guides(color = guide_legend(override.aes = list(size = 3), nrow = 4))
UMAP.SeuInt32 <-
  DimPlot(SeuInt3, group.by = "seurat_clusters", cols = colors_n, cells =  sample(Cells(SeuInt3)))
(UMAP.SeuInt31 +  UMAP.SeuInt32) &
  theme(legend.position = "bottom",
        legend.key.size = unit(0.1, "cm"))
# Plot QC graphics per cluster ####
#QC_plots(SeuInt3, "", colors_n)
```

### Color the UMAPs by library size:

The plots differ in how the scale is adjusted.

```{r UMAP_lbsz,  out.width= '100%', out.height= '75%', fig.asp=1.5}
thfp <-
  theme(plot.title = element_text(size = 9), legend.text = element_text(size = 4), axis.text = element_text(size = 4), axis.title = element_text(size = 4))
fp1 <-
  FeaturePlot(object = SeuInt3,
              features = "UMIs.p.cell",
              cells =  sample(Cells(SeuInt3))) + scale_colour_gradient(low = "lightblue", high = "darkblue") + 
  labs(title = "All; no lim on lbsz")+ thfp

fp2 <-
  FeaturePlot(object = SeuInt3,
              features = "UMIs.p.cell",
              cells =  sample(Cells(SeuInt3))) + scale_colour_gradient(
                limits = c(0, 10000),
                low = "lightblue",
                high = "darkblue"
              ) +  labs(title = "All; scale lim to 10000") + thfp

fp3 <-
  FeaturePlot(object = SeuInt3,
              features = "UMIs.p.cell",
              cells =  sample(Cells(SeuInt3))) + scale_colour_gradient(
                limits = c(0, 20000),
                low = "lightblue",
                high = "darkblue"
              ) + labs(title = "All; scale lim to 20000") +thfp

fp4 <-
  FeaturePlot(
    object = subset(SeuInt3, subset = Treat == "Ctrl"),
    features = "UMIs.p.cell",
    cells =  sample(Cells(subset(
      SeuInt3, subset = Treat == "Ctrl"
    )))
  ) + scale_colour_gradient(
    limits = c(0, 20000),
    low = "lightblue",
    high = "darkblue"
  ) + labs(title = "Ctrl; scale lim to 20000") + thfp

fp5 <-
  FeaturePlot(
    object = subset(SeuInt3, subset = Treat == "PD"),
    features = "UMIs.p.cell",
    cells =  sample(Cells(subset(
      SeuInt3, subset = Treat == "PD"
    )))
  ) + scale_colour_gradient(
    limits = c(0, 20000),
    low = "lightblue",
    high = "darkblue"
  ) + labs(title = "PD; scale lim to 20000") + thfp

fp6 <-
  FeaturePlot(
    object = subset(SeuInt3, subset = Treat == "Ctrl" &
                      Age == "1mo"),
    features = "UMIs.p.cell",
    cells =  sample(Cells(
      subset(SeuInt3, subset = Treat == "Ctrl" &
               Age == "1mo")
    ))
  ) + scale_colour_gradient(
    limits = c(0, 20000),
    low = "lightblue",
    high = "darkblue"
  ) + labs(title = "Ctrl 1mo; scale lim to 20000") + thfp

fp7 <-
  FeaturePlot(
    object = subset(SeuInt3, subset = Treat == "Ctrl" &
                      Age == "2mo"),
    features = "UMIs.p.cell",
    cells =  sample(Cells(
      subset(SeuInt3, subset = Treat == "Ctrl" &
               Age == "2mo")
    ))
  ) + scale_colour_gradient(
    limits = c(0, 20000),
    low = "lightblue",
    high = "darkblue"
  ) + labs(title = "Ctrl 2mo; scale lim to 20000") + thfp

plot_grid(fp1, fp2, fp3, fp4, fp5, fp6, fp7, NULL, ncol = 2)
```



```{r DataSummary2, include=T}
# Data summary
print("data summary integration")
data_summary <-
  data.frame(
    "value" = c(
      "date" = date,
      "path" = paste0(wd, FilePath, FileName),
      "object_name" = "SeuInt3",
      "filtered" = "no",
      "nr_cells" = dim(SeuInt3)[2],
      "demultiplexed" = "yes",
      "demux_positive.quantile" = 0.99,
      "norm.meth" =  "CLR",
      # Change this if needed
      HTO_classification,
      "integrated" = "yes",
      "ASW_batch" = round(ASW_batch, 4) ,
      "ASW_dims" = obj_dims,
      "nr_clusters" = length(levels(SeuInt3@meta.data$seurat_clusters))
    )
  ) %>% mutate("vars" = rownames(.), .before = value)  %>% `rownames<-`(NULL)
neigh_params <-
  SeuInt3$FindNeighbors.integrated.pca@params %>% unlist() %>% as.data.frame() %>% `colnames<-`("value")  %>% mutate("vars" = rownames(.), .before = value)  %>% `rownames<-`(NULL) %>% rbind(data.frame("vars" =  "neigh.params", "value" = NA), .)
clust_params <-
  SeuInt3$FindClusters@params %>% unlist() %>% as.data.frame() %>% `colnames<-`("value")  %>% mutate("vars" = rownames(.), .before = value)  %>% `rownames<-`(NULL) %>% rbind(data.frame("vars" =  "clust.params", "value" = NA), .)
data_summary <-
  rbind(data_summary, neigh_params) %>% rbind(clust_params)
data_summary
```

```{r DataSummary3, include=T}
## Data summary
print("data summary integrated filtered")
data_summary <-
  data.frame(
    "value" = c(
      "date" = date,
      "path" = paste0(wd, FilePath, FileName),
      "object_name" = "SeuInt",
      "filtered" = "yes",
      "UMIs.p.cell_lowthreshold" = UMIs.p.cell_lowthreshold,
      "UMIs.p.cell_highthreshold" = UMIs.p.cell_highthreshold,
      "percent.rb_threshold" = percent.rb_threshold,
      "nr_cells" = dim(SeuInt)[2],
      "demultiplexed" = "yes",
      "demux_positive.quantile" = 0.99,
      "norm.meth" =  "CLR",
      # Change this if needed
      HTO_classification,
      "integrated" = "yes",
      "ASW_batch" = round(ASW_batch, 4) ,
      "ASW_dims" = obj_dims,
      "nr_clusters" = length(levels(SeuInt@meta.data$seurat_clusters))
    )
  ) %>% mutate("vars" = rownames(.), .before = value)  %>% `rownames<-`(NULL)
neigh_params <-
  SeuInt$FindNeighbors.integrated.pca@params %>% unlist() %>% as.data.frame() %>% `colnames<-`("value")  %>% mutate("vars" = rownames(.), .before = value)  %>% `rownames<-`(NULL) %>% rbind(data.frame("vars" =  "neigh.params", "value" = NA), .)
clust_params <-
  SeuInt$FindClusters@params %>% unlist() %>% as.data.frame() %>% `colnames<-`("value")  %>% mutate("vars" = rownames(.), .before = value)  %>% `rownames<-`(NULL) %>% rbind(data.frame("vars" =  "clust.params", "value" = NA), .)
data_summary <-
  rbind(data_summary, neigh_params) %>% rbind(clust_params)
data_summary
```

```{r EGFP_hSNCA, include = T}
#Annotate which cell express KASH-EGFP, hSNCA-A53T and mCherry
print("get info of EGFP, hSNCA and mCherry positive cells")
DefaultAssay(SeuInt3) <- "RNA"
SeuInt3@meta.data$EGFP <-
  ifelse(GetAssayData(SeuInt3, slot = "counts")["EGFP-KASH", ] > 0, 1, 0) %>% as.factor()
SeuInt3@meta.data$hSNCA <-
  ifelse(GetAssayData(SeuInt3, slot = "counts")["hSNCA-A53T", ] > 0, 1, 0) %>% as.factor()
```

### Neuronal markers:

-   Slc6a5 = Noradrenaline Transporter (NET)

-   Th = Catalize L-DOPA production (precursor of Dopamine)

-   Dbh = Catalize Noradrenaline production from Dopamine

-   lightblue = GABAergic

-   lightgreen = Glutamatergic

-   brown = Cholinergic

-   violet = DopamineTransporter

-   red = AmineTranporter

-   darkbrown = Serotoninergic

-   orange = AmineDegradetion

```{r NeuronalMarkers,  out.width= '150%', out.height= '50%', fig.asp=1.2}
Dopaminergic <- c("Th", "Slc6a3") #"Th": Dopamine synthesis enzyme "Slc6a3": #DopamineTransporter
NoradrenalineTransporter <- "Slc6a5"
GABAergic <-
  c("Gad1", "Gad2") #"Gad1", "Gad2" : GABA synthesis enzymes
Glutamatergic <-
  c("Slc17a7", "Slc17a6", "Slc17a8") #"Slc17a7", "Slc17a6", "Slc17a8": Glutamate transporters
Cholinergic <-
  c("Slc18a3", "Slc5a7") #"Slc18a3", "Slc5a7": Acetylcholine transporters "Gm5741"?
Noradrenergic <- "Dbh" 
AmineTranporter <-
  c("Slc18a2", "Vat1") #"Slc18a2": Monoamine transporter (dopamine, norepinephrine, serotonin, and histamine) #"Vat1": Vesicle amine transporter
Serotoninergic <-
  c("Tph1", "Htr1a", "Htr1b", "Htr1f", "Htr1d",  "Htr2c") #"Tph1": Serotonin synthesis enzyme  #"Htr1a", "Htr1b", "Htr1f", "Htr1d",  "Htr2c": Serotonin transporter
AmineDegradetion <-
  c("Comt", "Maoa", "Maob") #"Comt", "Maoa", "Maob": Amine degradetion

# Plot neuronal markers
DefaultAssay(SeuInt3) <-
  "RNA" # Th gene is not a variable feature in the INTE assay
th <-
  theme(
    text = element_text(size = 8),
    axis.text = element_text(size = 5),
    plot.title = element_text(size = 10, face = "bold")
  )
plot1 <-
  FeaturePlot(
    SeuInt3,
    features = c(Dopaminergic, Noradrenergic, "EGFP-KASH"),
    pt.size = 0.00001,
    order =  T,
    ncol = length(c(Dopaminergic, Noradrenergic, "EGFP-KASH"))
  ) & th #normalized counts
SeuInt3@meta.data$Thg <-
  ifelse(GetAssayData(SeuInt3, slot = "counts")["Th", ] > 0, 1, 0) %>% as.factor()
SeuInt3@meta.data$Slc6a3g <-
  ifelse(GetAssayData(SeuInt3, slot = "counts")["Slc6a3", ] > 0, 1, 0) %>% as.factor()
SeuInt3@meta.data$Dbhg <-
  ifelse(GetAssayData(SeuInt3, slot = "counts")["Dbh", ] > 0, 1, 0) %>% as.factor()
plot2 <- DimPlot(SeuInt3, group.by = "Thg", order = T) + th
plot3 <- DimPlot(SeuInt3, group.by = "Slc6a3g", order = T) + th
plot4 <- DimPlot(SeuInt3, group.by = "Dbhg", order = T) + th
layout1 <- "
ABC
DFG
HIJ
KLM
"
plot5 <-
  FeaturePlotMultiple2(
    SeuInt3,
    feature = c(
      GABAergic,
      Glutamatergic,
      Cholinergic,
      Noradrenergic,
      AmineTranporter
    ),
    color = rep(
      c("deepskyblue3", "limegreen", "chocolate2", "violetred2", "red"),
      times = c(2, 3, 2, 1, 2)
    ),
    ord = T,
    alf = 0.5,
    minlim = "q1"
  )
plot5 <-
  wrap_plots(plot5, guides = "keep", design = layout1) # Slc18a3, Slc17a7, Slc17a8 very low expression
plot6 <-
  FeaturePlotMultiple2(
    SeuInt3,
    feature = c(Serotoninergic, AmineDegradetion),
    color = rep(c("#6e2c00", "darkorange"), times = c(6, 3)),
    ord = T,
    alf = 0.5,
    minlim = "q1"
  )
plot6 <-
  wrap_plots(plot6, guides = "keep", design = layout1) # Tph1, Maob very low expression
palfa <- plot_grid(plot1 , plot_grid(plot2, plot3, plot4, NULL, nrow = 1), ncol = 1)
pbeta <- plot_grid(plot5, plot_grid(plot6, NULL, ncol = 1, rel_heights = c(8, 1)), nrow = 1)
plot_grid(palfa, pbeta, ncol = 1, rel_heights = c(0.5, 1)) 
```

```{r}
# Percentage_exp <- apply(GetAssayData(SeuInt, slot = "counts"), 1, function(x){x*100/SeuInt$UMIs.p.cell}) %>% as.data.frame()
# pivot_longer(Percentage_exp, cols)

```


## The normalized expression of transgenes markers

The enrichment was calculated without taking into account the replicates, with fisher exact test. The odds ratio of the `fisher.test()` is estimated by an MLE, based on a Non Centered Hypergeometric distribution supported on the observed marginal probabilities and the p-value is obtained by getting the probability in which a number related to the first element of the matrix is located in a shifted distribution to be "centered" in an OR of 1. For more info try to intepret the source code or its documentation references [here](https://github.com/SurajGupta/r-source/blob/master/src/library/stats/R/fisher.test.R)

```{r TransgeneExpression}
# Plot the UMAP colored by EGFP and hSNCA expressin and add a UMAP of clusters by the side and the enrichment of EGFP+hSNCA+ cells on PD Treatment
# Plot UMAP colored by EGFP+ OR EGFP- - continuous scale
SeuInt3@meta.data$Genes <- ifelse(SeuInt3@meta.data$EGFP=="1"&SeuInt3@meta.data$hSNCA =="1", "both",
                                    ifelse(SeuInt3@meta.data$EGFP=="0"&SeuInt3@meta.data$hSNCA =="0", "none",
                                           ifelse(SeuInt3@meta.data$EGFP == "1", "KASH-EGFP", "hSNCA-A53T"))) %>% factor(levels = c("none", "KASH-EGFP", "hSNCA-A53T", "both"))


# text <-
#   ggplot(data.frame(x = 1, y = 1)) + geom_text(
#     size = 3,
#     label = paste0("EGFP+SNCA+ enrichment: ", paste(label2analyze_EGFP_hSNCA_FisherTest, collapse = "")),
#     x = 0,
#     y = 0,
#     hjust = 0) + theme_void()
SeuInt3_PD <- subset(SeuInt3, subset = Treat == "PD")

text <- ""
DefaultAssay(SeuInt3_PD) <- "integrated"
UMAP_grid <- UMAPs_for_grid(SeuInt3_PD, "Clusters")
UMAP_PD <- plot_grid(plot_grid(UMAP_grid[["UMAP4_split"]], UMAP_grid[["UMAPa_edit"]], ncol = 2), plot_grid(NULL, UMAP_grid[["UMAP4_legend"]], ncol = 2, rel_widths = c(0.1, 3)), text, ncol = 1, nrow = 3, rel_heights = c(4,0.2,0.7))
UMAP_PD
```

#### Fisher exact test

First try with the fisher exact test, pooling all cells of each treatment together (no condition DS stratification). Part of the interpretation of the odds ratio can be found [here](https://stats.stackexchange.com/questions/211487/interpretation-of-the-fisher-exact-test) and a short discussion about its p-value can be found  [here](https://www.scalestatistics.com/fishers-exact-test.html)
```{r}
# Calculate the global- or cluster-specific- contingency table of the expression of hSNCA and EGFP

# Obtain the global contingency
global_contingency <- select(SeuInt3_PD@meta.data, "seurat_clusters", "hSNCA", "EGFP") %>% do(data.frame(      
    EGFP_SNCA = EGFP_SNCA(.),
      EGFP = EGFP(.),
      SNCA = SNCA(.),
      Neith = Neith(.)
    ))

# Obtain the cluster contingency
cluster_contingency <- select(SeuInt3_PD@meta.data, "seurat_clusters", "hSNCA", "EGFP") %>% group_by(seurat_clusters) %>%
  do(data.frame(      
    EGFP_SNCA = EGFP_SNCA(.),
      EGFP = EGFP(.),
      SNCA = SNCA(.),
      Neith = Neith(.)
    ))

# Obtain the "rest of cells" contingency for each cluster
rest_contingency <- apply(
  as.matrix(cluster_contingency[,-1]), #Take the values of EGFP and hSNCA presence per cluster
  1, 
  function(x) -1*x + as.vector(global_contingency[1,])) %>% #And get the global EGFP and hSNCA presence minus the presence in each of the clusters
  do.call(what = rbind) %>% as.data.frame()
rest_contingency$seurat_clusters <- cluster_contingency$seurat_clusters

# Obtain the contingency table for the presence/absence of hSNCA (alone or together with EGFP)
# for the cells of each cluster and the rest of the cells of the data set

hSNCA_contingency <- data.frame(
  yes_cluster = cluster_contingency$SNCA + cluster_contingency$EGFP_SNCA,
  no_cluster = cluster_contingency$EGFP + cluster_contingency$Neith,
  yes_rest_cluster = rest_contingency$SNCA + rest_contingency$EGFP_SNCA,
  no_rest_cluster = rest_contingency$EGFP + rest_contingency$Neith
  )

# Obtain the contingency table for the presence/absence of hSNCA with EGFP
# for the cells of each cluster and the rest of the cells of the data set

EGFP_hSNCA_contingency <- data.frame(
  yes_cluster = cluster_contingency$EGFP_SNCA,
  no_cluster = cluster_contingency$EGFP + cluster_contingency$Neith +
    cluster_contingency$SNCA,
  yes_rest_cluster = rest_contingency$EGFP_SNCA,
  no_rest_cluster = rest_contingency$EGFP + rest_contingency$Neith +
    rest_contingency$SNCA
  )

# Calculate the enrichment of hSNCA for each cluster compared to rest of cells
hSNCA_FisherTest<-apply(as.matrix(hSNCA_contingency), 1, fisher)
colnames(hSNCA_FisherTest) <- rest_contingency$seurat_clusters
hSNCA_FisherTest<- rbind(hSNCA_FisherTest, "FDR" = p.adjust(hSNCA_FisherTest["p_val",], "BH"))

# Calculate the enrichment of hSNCA with EGFP for each cluster compared to rest of cells
EGFP_hSNCA_FisherTest<-apply(as.matrix(EGFP_hSNCA_contingency), 1, fisher)
colnames(EGFP_hSNCA_FisherTest) <- rest_contingency$seurat_clusters
EGFP_hSNCA_FisherTest<- rbind(EGFP_hSNCA_FisherTest, "FDR" = p.adjust(EGFP_hSNCA_FisherTest["p_val",], "BH"))

# Calculate the enrichment of hSNCA with EGFP but instead of taking as control the rest of the cells as it is, I will "shuffle" the tags for EGFP and SNCA, so that the ocurrence of both genes will be drived randomly (is a way to see if the cluster itself has enriched the co-mark of hSNCA and EGFP, unbiased big cluster enrichment on it).

# Obtain parameters for shuffled dataset
shuffled_rest_param <- data.frame(EGFP = rest_contingency$EGFP_SNCA + rest_contingency$EGFP, SNCA = rest_contingency$EGFP_SNCA + rest_contingency$SNCA, Neith = rest_contingency$Neith) %>% mutate(Total = EGFP + SNCA + Neith, Prop_EGFP = EGFP/Total, Prop_SNCA = SNCA/Total, Prop_Neith = Neith/Total)
list_shuffled_rest_param <- split(shuffled_rest_param[4:7],seq(nrow(shuffled_rest_param[4:7])))   
  
# Create random numbers of EGFP, SNCA, both and Neith based on parameters
shuffled_rest_contingency <- lapply(list_shuffled_rest_param, function(x) create_random(x[, "Total"], x[, "Prop_EGFP"], x[, "Prop_SNCA"], x[, "Prop_Neith"])) %>% do.call(what = rbind) %>% as.data.frame()

# Make contingency table
shuff_EGFP_hSNCA_contingency <- data.frame(
  yes_cluster = cluster_contingency$EGFP_SNCA,
  no_cluster = cluster_contingency$EGFP + cluster_contingency$Neith +
    cluster_contingency$SNCA,
  yes_rest_cluster = shuffled_rest_contingency$EGFP_SNCA,
  no_rest_cluster = shuffled_rest_contingency$EGFP + shuffled_rest_contingency$Neith +
    shuffled_rest_contingency$SNCA
  )

# Calculate the enrichment of hSNCA with EGFP for each cluster compared to random cells
shuff_EGFP_hSNCA_FisherTest<-apply(as.matrix(shuff_EGFP_hSNCA_contingency), 1, fisher)
colnames(shuff_EGFP_hSNCA_FisherTest) <- rest_contingency$seurat_clusters
shuff_EGFP_hSNCA_FisherTest<- rbind(shuff_EGFP_hSNCA_FisherTest, "FDR" = p.adjust(shuff_EGFP_hSNCA_FisherTest["p_val",], "BH"))
## Pero esta bien hacer esto una sola vez? Lo deberia repetir. No serian datos imputados. Si lo repito como combino los odds.ratios y los p-values.

# Obtain the contingency just for the combination of both variables for each cluster. Not very informative for our aims since we already jnow that, depite the results, all SNCA+ cells have been sorted, so theyr epress EGFP at some point (despite the detectable expression).
contingency <- select(SeuInt3_PD@meta.data, "seurat_clusters", "hSNCA", "EGFP") %>% group_by(seurat_clusters) %>%
  do(data.frame(      
    EGFP_SNCA = EGFP_SNCA(.),
      EGFP = EGFP(.),
      SNCA = SNCA(.),
      Neith = Neith(.)
    ))
FisherTest<-apply(as.matrix(contingency[, -1]), 1, fisher)
colnames(FisherTest) <- contingency$seurat_clusters
FisherTest<- rbind(FisherTest, "FDR" = p.adjust(FisherTest["p_val",], "BH"))

# Add a text for the enrichment
label2analyze_FisherTest <- label2analyze(FisherTest)
label2analyze_shuff_EGFP_hSNCA_FisherTest<- label2analyze(shuff_EGFP_hSNCA_FisherTest)
label2analyze_EGFP_hSNCA_FisherTest <- label2analyze(EGFP_hSNCA_FisherTest)
label2analyze_hSNCA_FisherTest <- label2analyze(hSNCA_FisherTest)
```

The enrichment are:

EGFP+SNCA+ enrichment: 

```{r}
label2analyze_EGFP_hSNCA_FisherTest
```

shuffled EGFP+SNCA+ enrichment:
```{r}
label2analyze_shuff_EGFP_hSNCA_FisherTest
```

SNCA+ enrichment:
```{r}
label2analyze_hSNCA_FisherTest
```


```{r}
# SeuInt3@meta.data$Genes <-
#   ifelse(
#     SeuInt3@meta.data$EGFP == "1" & SeuInt3@meta.data$hSNCA == "1",
#     "both",
#     ifelse(
#       SeuInt3@meta.data$EGFP == "0" & SeuInt3@meta.data$hSNCA == "0",
#       "none",
#       ifelse(SeuInt3@meta.data$EGFP == "1", "KASH-EGFP", "hSNCA-A53T")
#     )
#   ) %>% factor(levels = c("none", "KASH-EGFP", "hSNCA-A53T", "both"))
# SeuInt3_PD <- subset(SeuInt3, subset = Treat == "PD")
# 
# # Add a text with the EGFP+hSNCA+ cells in PD dataset
# contingency <-
#   select(SeuInt3_PD@meta.data, "seurat_clusters", "hSNCA", "EGFP") %>% group_by(seurat_clusters) %>%
#   do(data.frame(
#     Neith = Neith(.),
#     SNCA = SNCA(.),
#     EGFP = EGFP(.),
#     EGFP_SNCA = EGFP_SNCA(.)
#   ))
# FisherTest <- apply(as.matrix(contingency[, -1]), 1, fisher)
# colnames(FisherTest) <- contingency$seurat_clusters
# FisherTest <-
#   rbind(FisherTest, "FDR" = p.adjust(FisherTest["p_val",], "BH"))
# label2analyze <- c()
#   for (i in 1:length(clusters2analyze)){
#     if (i != 4){
#       label2analyze[i] <- paste0("Cluster ", clusters2analyze[i], " : ", round(FisherTest["enrich", clusters2analyze[i]], 2), " (p-adj:", format(FisherTest["FDR", clusters2analyze[i]], digits=3), ") ; ")
#     } else {
#       label2analyze[i] <- paste0("Cluster ", clusters2analyze[i], " : ", round(FisherTest["enrich", clusters2analyze[i]], 2), " (p-adj:", format(FisherTest["FDR", clusters2analyze[i]], digits=3), ") ;\n")
#     }
#   }
# text <-
#   ggplot(data.frame(x = 1, y = 1)) + geom_text(
#     size = 3,
#     label = paste0("EGFP+SNCA+ enrichment: ", paste(label2analyze, collapse = "")),
#     x = 0,
#     y = 0.5,
#     hjust = 0) + theme_void()
# DefaultAssay(SeuInt3_PD) <- "integrated"
# UMAP_grid <- UMAPs_for_grid(SeuInt3_PD, "Clusters")
# UMAP_PD <-
#   plot_grid(
#     plot_grid(UMAP_grid[["UMAP4_split"]], UMAP_grid[["UMAPa_edit"]], ncol = 2),
#     plot_grid(
#       NULL,
#       UMAP_grid[["UMAP4_legend"]],
#       ncol = 2,
#       rel_widths = c(0.1, 3)
#     ),
#     text,
#     ncol = 1,
#     nrow = 3,
#     rel_heights = c(4, 0.2, 0.5)
#   )
# UMAP_PD
```

```{r, include=T}
rm(SeuInt3_PD); gc()
```

#### How many cells for each cluster?

```{r}
table(SeuInt3$seurat_clusters, SeuInt3$Treat, SeuInt3$Age)
# SeuInt3 <- subset(SeuInt3, subset = seurat_clusters != "20")
# SeuInt3$seurat_clusters <- factor(SeuInt3$seurat_clusters, levels = as.character(c(0:19)))
```


#### Analyze for each month:

### UMAP with \# of cells per cluster:

```{r Nr.cells.month, out.width= '100%'}
#### For each month ####
SeuInt3_sub <- list()
ncells <- list()
for (i in c("1mo", "2mo")) {
  SeuInt3_sub[[i]] <- subset(SeuInt3, subset = Age == i)
  ## Nr. cells per cluster ####
  # Generate Labels of clusters and calculate number of cells of each cluster
  # Number of PD cells in whole dataset and on each cluster
  ncellsPD <-
    select(SeuInt3_sub[[i]]@meta.data, "seurat_clusters", "Treat") %>% group_by(seurat_clusters, Treat) %>%
    summarize(n.cells = n()) %>% filter(Treat == "PD")
  tot.PD <- sum(ncellsPD$n.cells)
  # Number of Ctrl cells in whole dataset and on each cluster
  ncellsCtrl <-
    select(SeuInt3_sub[[i]]@meta.data, "seurat_clusters", "Treat") %>% group_by(seurat_clusters, Treat) %>%
    summarize(n.cells = n()) %>% filter(Treat == "Ctrl")
  tot.Ctrl <- sum(ncellsCtrl$n.cells)
  # DataFrame with: dif.prop = Difference of PD and Ctrl percentage on each cluster; diff.abs = diff.prop but in absolute value;
  # PD.text = label of number of PD cells on each cluster; Ctrl.text = label of number of Ctrl cells on each cluster
  ncells_i <-
    left_join(ncellsPD, ncellsCtrl, by = "seurat_clusters") %>% mutate(
      dif.prop = round(100 * (n.cells.x - n.cells.y) / n.cells.y, 0),
      dif.abs = paste0(seurat_clusters, " (", n.cells.x, ":", n.cells.y, ")"),
      `PD.text` = paste0(seurat_clusters, " (", n.cells.x, ")"),
      `Ctrl.text` = paste0(seurat_clusters, " (", n.cells.y, ")")
    ) %>% select(seurat_clusters, dif.prop, dif.abs, `PD.text`, `Ctrl.text`)
  
  # Total cells on each cluster
  totcells <-
    select(SeuInt3_sub[[i]]@meta.data, "seurat_clusters", "Treat") %>% group_by(seurat_clusters) %>%
    summarize(tot.cells = n())
  # DataFrame with previous info plus: brak.text = label of total cells and dif.prop; Int.text = label of percentage of cells from whole dataset
  ncells[[i]] <-
    left_join(ncells_i, totcells) %>% mutate(
      brak.text = paste0(seurat_clusters, " (", tot.cells, ": ", dif.prop, " %)"),
      Int.text = paste0(seurat_clusters, " (", round(100 *
                                                       tot.cells / dim(SeuInt3_sub[[i]])[2], 4), "%)")
    )
  
  ## Plot UMAP of integrated data ####
  # Plot by clusters with #Cells: PD % Dif and by Treatment
  ClusterLabels <-  ncells[[i]]$brak.text
  ClusterBreaks <- ncells[[i]]$seurat_clusters
  title_text <-
    paste0(
      i,
      " > ",
      UMIs.p.cell_lowthreshold,
      " <",
      UMIs.p.cell_highthreshold,
      " UMI < ",
      percent.rb_threshold,
      " per.rb Matrix (",
      dim(SeuInt3_sub[[i]])[2],
      " Cells)"
    )
  subtitle_text <- "r=0.1; k=10; dim=50; Louvain"
  color_a <- "Clusters (#Cells: PD % Dif)"
  color_b <- "Treatment"
  
  UMAP1 <-
    UMAP_by_clust(
      SeuInt3_sub[[i]],
      ClusterBreaks,
      ClusterLabels,
      0.5,
      title_text,
      subtitle_text,
      color_a,
      color_b,
      do_label = F,
      n_color = colors_n
    ) #Ojo que hay un aparte del titulo que se superpone con otra
  UMAP1a <- UMAP1[["UMAPa"]] + UMAP1[["UMAPb"]]
  print(UMAP1a)
}

for (i in c("1mo", "2mo")) {
  ## Plot UMAP of integrated data ####
  # Plot by clusters with #Cells: PD % Dif and by Treatment
  ClusterLabels <-  ncells[[i]]$brak.text
  ClusterBreaks <- ncells[[i]]$seurat_clusters
  title_text <-
    paste0(
      i,
      " > ",
      UMIs.p.cell_lowthreshold,
      " <",
      UMIs.p.cell_highthreshold,
      " UMI < ",
      percent.rb_threshold,
      " per.rb Matrix (",
      dim(SeuInt3_sub[[i]])[2],
      " Cells)"
    )
  subtitle_text <- "r=0.1; k=10; dim=50; Louvain"
  color_a <- "Clusters (#Cells: PD % Dif)"
  color_b <- "Treatment"
  
  UMAP1 <-
    UMAP_by_clust(
      SeuInt3_sub[[i]],
      ClusterBreaks,
      ClusterLabels,
      0.5,
      title_text,
      subtitle_text,
      color_a,
      color_b,
      do_label = F,
      n_color = colors_n
    ) #Ojo que hay un aparte del titulo que se superpone con otra
  # Plot by clusters with #Cells: PD % Dif and by Treatment
  UMAP1b <-
    (
      LabelClusters(
        UMAP1[["UMAPa"]],
        id = "ident",
        size = 4,
        repel = F
      ) + theme(legend.position = "none")
    ) +  UMAP1[["UMAPb"]]
  #print(UMAP1b)
}
```

### UMAP by clusters with #Cells PD:Ctrl, splitted by treatment:

```{r PD:Ctrl,  out.width= '85%', out.height= '25%', fig.asp=0.5}
for (i in c("1mo", "2mo")) {
  ClusterLabels <-  ncells[[i]]$dif.abs
  ClusterBreaks <- ncells[[i]]$seurat_clusters
  UMAP2 <-
    DimPlot(
      SeuInt3_sub[[i]],
      reduction = "umap",
      label = T,
      pt.size = 0.5,
      split.by = "Treat",
      cols = colors_n
    ) +
scale_colour_manual(values = colors_n, breaks = ClusterBreaks, labels = ClusterLabels) +
    labs(
      title = paste0(
        i,
        " > ",
        UMIs.p.cell_lowthreshold,
        " <",
        UMIs.p.cell_highthreshold,
        " UMI < ",
        percent.rb_threshold,
        " per.rb Matrix (",
        dim(SeuInt3_sub[[i]])[2],
        " Cells)"
      ),
      subtitle = "r=0.1; k=10; dim=50; Louvain",
      color = "Clusters (#Cells PD:Ctrl)"
    ) +
    theme_classic(base_size = 10) +
    theme(legend.text = element_text(size = 6)) +
    guides(color = guide_legend(override.aes = list(size = 1), ncol = 2))
  print(UMAP2)
}
```

### QC graphics:

```{r QC,  out.width= '100%'}
for (i in c("1mo", "2mo")) {
  DefaultAssay(SeuInt3_sub[[i]]) <- "RNA"
  QC_plot_Int <- QC_plots(SeuInt3_sub[[i]], i, colors_n)
  print(QC_plot_Int)
}
```

```{r KASH-EGFP+, include=T, out.width= '50%', out.height= '50%'}
clusters_EGFP <- list()
for (i in c("1mo", "2mo")) {
  clusters_EGFP_tot <-
    select(SeuInt3_sub[[i]]@meta.data, "seurat_clusters", "EGFP", "Treat") %>% group_by(seurat_clusters, Treat) %>%
    summarize(
      sum.EGFP = sum(grepl(1, EGFP)),
      tot = n(),
      per.EGFP.treat = 100 * (sum.EGFP / tot)
    ) %>% ungroup()
  # DataFrame with previous info, adding:
  # perc = percentage of EGFP+ cells from all EGFP+ cells in that treatment
  # perc_text = the text of percentage of EGFP+ cells for cluster 0 and 4.
  clusters_EGFP_tot <-
    group_by(clusters_EGFP_tot, Treat) %>% summarize(tot.EGFP.orig = sum(sum.EGFP)) %>% ungroup() %>%
    left_join(clusters_EGFP_tot) %>%
    mutate(
      perc = sum.EGFP / tot.EGFP.orig,
      perc_text = ifelse(
        seurat_clusters %in% text4stackbarEGFP,
        paste0(round(100 * sum.EGFP / tot.EGFP.orig, 2), "%"),
        ""
      )
    )
  # per.EGFP.tot: percentage of EGFP+ cells from total cells without considering the treatment
  clusters_EGFP[[i]] <-
    group_by(clusters_EGFP_tot, seurat_clusters) %>% summarize(
      tot.orig = sum(tot),
      per.EGFP.tot = 100 * sum(sum.EGFP) / tot.orig,
      no.EGFP = ifelse(sum(sum.EGFP) == 0, T, F)
    ) %>% ungroup %>% left_join(clusters_EGFP_tot)
  # Plot the perc as a stacked ber. Each bar is a treatment and clusters for each treatment are stacked.
}
# Observations:
# How to get the clusters that belong to the NA neurons of the LC?
# - By the percentage of EGFP on the control treatment?
#   - How to determine the threshold of the percentage to assign that cluster to a NA-LC cluster?
#   - Do I have to filter out clusters that have too few cells depite having a high EGFP percentage. How much is few?
#   - Do I have to consider an EGFP cluster when the I have a high percentage only for the PD treatment but not for the Ctrl?
#   - Can I do all this with a assuming a Bernoulli probability? Does this make sense depite samples are not independent?
# - By enrichment of LC genes by a GSVA analysis?
# - By another classification algorithm for single cells?
```

### Number and percentage of KASH-EGFP+ cells:

```{r,  out.width= '50%', out.height= '50%'}
for (i in c("1mo", "2mo")) {
  Stacked_bar <-
    ggplot(clusters_EGFP[[i]],
           aes(fill = seurat_clusters, x = Treat, y = perc)) + #aes(fill=seurat_clusters, y=perc, x=Treat)
    geom_bar(position = "stack",
             stat = "identity",
             color = "white") + #position="fill"
    # scale_fill_manual(values = colors_n[as.character(c(0:20))]) +
    geom_text(aes(label = perc_text),
              position = position_stack(vjust = 0.5),
              size = 5) +
    labs(
      title = i,
      fill = "Clusters",
      y = "Number of EGFP cells",
      x = "Treatment"
    ) +
    my_theme +
    scale_fill_manual(values = colors_n, breaks = filter(clusters_EGFP[[i]], no.EGFP == F)$seurat_clusters) +
    scale_y_continuous(breaks = seq(0, 300, by = 10), labels = insert_minor(seq(0, 300, by =
                                                                                  50), 4)) +
    geom_area(
      aes(x = c("PD" = 1.55, "Ctrl" = 1.45)[Treat]),
      position = "stack",
      colour = "white",
      alpha = 0.5,
      #position="fill"
      outline.type = "both"
    )
  print(Stacked_bar)
  ## Horizontal stacked bar of number and percentage of EGFP+ cells
  # Stacked_bar_horiz <- Stacked_bar + coord_flip()
  # print(Stacked_bar_horiz)
}
```

### Number and percentage of KASH-EGFP+ cells:

```{r barKASH-EGFP+,  out.width= '50%', out.height= '50%'}
for (i in c("1mo", "2mo")) {
  Stacked_bar <-
    ggplot(clusters_EGFP[[i]],
           aes(fill = seurat_clusters, x = Treat, y = sum.EGFP)) + #aes(fill=seurat_clusters, y=perc, x=Treat)
    geom_bar(position = "stack",
             stat = "identity",
             color = "white") + #position="fill"
    # scale_fill_manual(values = colors_n[as.character(c(0:20))]) +
    geom_text(aes(label = perc_text),
              position = position_stack(vjust = 0.5),
              size = 5) +
    labs(
      title = i,
      fill = "Clusters",
      y = "Number of EGFP cells",
      x = "Treatment"
    ) +
    my_theme +
    scale_fill_manual(values = colors_n, breaks = filter(clusters_EGFP[[i]], no.EGFP == F)$seurat_clusters) +
    scale_y_continuous(breaks = seq(0, 300, by = 10), labels = insert_minor(seq(0, 300, by =
                                                                                  50), 4)) +
    geom_area(
      aes(x = c("PD" = 1.55, "Ctrl" = 1.45)[Treat]),
      position = "stack",
      colour = "white",
      alpha = 0.5,
      #position="fill"
      outline.type = "both"
    )
  print(Stacked_bar)
  ## Horizontal stacked bar of number and percentage of EGFP+ cells
  # Stacked_bar_horiz <- Stacked_bar + coord_flip()
  # print(Stacked_bar_horiz)
}
```



### Number and percentage of cells of all cells (not only KASH-EGFP+):

```{r,  out.width= '50%', out.height= '50%'}
clusters_EGFP <- list()
for (i in c("1mo", "2mo")) {
  clusters_EGFP[[i]] <-
    select(SeuInt3_sub[[i]]@meta.data, "seurat_clusters", "Treat") %>% group_by(seurat_clusters, Treat) %>%
    summarize(tot = n()) %>% ungroup()
  # # Plot the perc as a stacked ber. Each bar is a treatment and clusters for each treatment are stacked.
  clusters_EGFP[[i]] <-
    group_by(clusters_EGFP[[i]], Treat) %>% summarize(TotTreat = sum(tot)) %>% ungroup %>% left_join(clusters_EGFP[[i]]) %>%
    mutate(
      percTot = tot / TotTreat,
      Int.text =  ifelse(
        seurat_clusters %in% text4stackbar,
        paste0(round(100 * percTot, 2), " %"),
        " "
      )
    )
  
  Stacked_bar <-
    ggplot(clusters_EGFP[[i]], aes(fill = seurat_clusters, x = Treat, y = percTot)) + #aes(fill=seurat_clusters, y=perc, x=Treat)
    geom_bar(position = "stack",
             stat = "identity",
             color = "white") + #position="fill"
    scale_fill_manual(values = colors_n) + #[as.character(c(0:19))]
    geom_text(aes(label = Int.text),
              position = position_stack(vjust = 0.5),
              size = 5) +
    labs(
      title = i,
      fill = "Clusters",
      y = "Number of cells",
      x = "Treatment"
    ) +
    my_theme +
    #scale_fill_discrete(breaks= filter(clusters_EGFP_tot, no.EGFP==F)$seurat_clusters) +
    scale_y_continuous(breaks = seq(0, 3000, by = 20),
                       labels = insert_minor(seq(0, 3000, by = 100), 4)) +
    geom_area(
      aes(x = c("PD" = 1.55, "Ctrl" = 1.45)[Treat]),
      position = "stack",
      colour = "white",
      alpha = 0.5,
      #position="fill"
      outline.type = "both"
    )
  print(Stacked_bar)
}
```

### Number and percentage of cells of all cells (not only KASH-EGFP+):

```{r barAll,  out.width= '50%', out.height= '50%'}
clusters_EGFP <- list()
for (i in c("1mo", "2mo")) {
  clusters_EGFP[[i]] <-
    select(SeuInt3_sub[[i]]@meta.data, "seurat_clusters", "Treat") %>% group_by(seurat_clusters, Treat) %>%
    summarize(tot = n()) %>% ungroup()
  # # Plot the perc as a stacked ber. Each bar is a treatment and clusters for each treatment are stacked.
  clusters_EGFP[[i]] <-
    group_by(clusters_EGFP[[i]], Treat) %>% summarize(TotTreat = sum(tot)) %>% ungroup %>% left_join(clusters_EGFP[[i]]) %>%
    mutate(
      percTot = tot / TotTreat,
      Int.text =  ifelse(
        seurat_clusters %in% text4stackbar,
        paste0(round(100 * percTot, 2), " %"),
        " "
      )
    )
  
  Stacked_bar <-
    ggplot(clusters_EGFP[[i]], aes(fill = seurat_clusters, x = Treat, y = tot)) + #aes(fill=seurat_clusters, y=perc, x=Treat)
    geom_bar(position = "stack",
             stat = "identity",
             color = "white") + #position="fill"
    scale_fill_manual(values = colors_n) + #[as.character(c(0:19))]
    geom_text(aes(label = Int.text),
              position = position_stack(vjust = 0.5),
              size = 5) +
    labs(
      title = i,
      fill = "Clusters",
      y = "Number of cells",
      x = "Treatment"
    ) +
    my_theme +
    # scale_fill_discrete(values = colors_n, breaks= filter(clusters_EGFP_tot, no.EGFP==F)$seurat_clusters) +
    scale_y_continuous(breaks = seq(0, 3000, by = 20),
                       labels = insert_minor(seq(0, 3000, by = 100), 4)) +
    geom_area(
      aes(x = c("PD" = 1.55, "Ctrl" = 1.45)[Treat]),
      position = "stack",
      colour = "white",
      alpha = 0.5,
      #position="fill"
      outline.type = "both"
    )
  print(Stacked_bar)
}
```

### Violin plots of expression of "EGFP-KASH" per treatment

```{r Vln.month,  out.width= '50%', out.height= '50%', fig.asp=0.75}
for (i in c("1mo", "2mo")) {
  VlnPlot1 <-
    VlnPlot(
      SeuInt3_sub[[i]],
      features = c("EGFP-KASH"),
      split.by = "Treat",
      split.plot = T,
      log = T
    ) + labs(title = i)
  print(VlnPlot1)
}
```

### UMAP with the enrichment of EGFP+hSNCA+ cells:

```{r EGFP+hSNCA+,  out.height= '110%'}
clusters2analyze <- c("0", "6", "7", "8", "18")
for (i in c("1mo", "2mo")) {
  SeuInt3_PD <- subset(SeuInt3_sub[[i]], subset = Treat == "PD")
  # Add a text with the EGFP+hSNCA+ cells in PD dataset
  global_contingency <- select(SeuInt3_PD@meta.data, "seurat_clusters", "hSNCA", "EGFP") %>% do(data.frame(      
    EGFP_SNCA = EGFP_SNCA(.),
      EGFP = EGFP(.),
      SNCA = SNCA(.),
      Neith = Neith(.)
    ))

# Obtain the cluster contingency
cluster_contingency <- select(SeuInt3_PD@meta.data, "seurat_clusters", "hSNCA", "EGFP") %>% group_by(seurat_clusters) %>%
  do(data.frame(      
    EGFP_SNCA = EGFP_SNCA(.),
      EGFP = EGFP(.),
      SNCA = SNCA(.),
      Neith = Neith(.)
    ))

# Obtain the "rest of cells" contingency for each cluster
rest_contingency <- apply(as.matrix(cluster_contingency[,-1]),  1, 
  function(x) -1*x + as.vector(global_contingency[1,])) %>%
  do.call(what = rbind) %>% as.data.frame()
rest_contingency$seurat_clusters <- cluster_contingency$seurat_clusters
  
  EGFP_hSNCA_contingency <- data.frame(
  yes_cluster = cluster_contingency$EGFP_SNCA,
  no_cluster = cluster_contingency$EGFP + cluster_contingency$Neith +
    cluster_contingency$SNCA,
  yes_rest_cluster = rest_contingency$EGFP_SNCA,
  no_rest_cluster = rest_contingency$EGFP + rest_contingency$Neith +
    rest_contingency$SNCA
  )
  
  # Calculate the enrichment of hSNCA with EGFP for each cluster compared to rest of cells
EGFP_hSNCA_FisherTest<-apply(as.matrix(EGFP_hSNCA_contingency), 1, fisher)
colnames(EGFP_hSNCA_FisherTest) <- rest_contingency$seurat_clusters
EGFP_hSNCA_FisherTest<- rbind(EGFP_hSNCA_FisherTest, "FDR" = p.adjust(EGFP_hSNCA_FisherTest["p_val",], "BH"))

label2analyze <- c()
  for (m in 1:length(clusters2analyze)){
    if (m != 4){
      label2analyze[m] <- paste0("Cluster ", clusters2analyze[m], " = ", round(EGFP_hSNCA_FisherTest["enrich", clusters2analyze[m]], 2), " (p-adj = ", format(EGFP_hSNCA_FisherTest["FDR", clusters2analyze[m]], digits=3), "); ")
    } else {
      label2analyze[m] <- paste0("Cluster ", clusters2analyze[m], " = ", round(EGFP_hSNCA_FisherTest["enrich", clusters2analyze[m]], 2), " (p-adj = ", format(EGFP_hSNCA_FisherTest["FDR", clusters2analyze[m]], digits=3), ");\n")
    }
  }
text <-
  ggplot(data.frame(x = 1, y = 1)) + geom_text(
    size = 3,
    label = paste0("EGFP+SNCA+ enrichment: ", paste(label2analyze, collapse = "")),
    x = 0,
    y = 0.5,
    hjust = 0) + theme_void()  
DefaultAssay(SeuInt3_PD) <- "integrated"
  UMAP_grid <- UMAPs_for_grid(SeuInt3_PD, "Clusters")
  UMAP5 <-
    plot_grid(
      ggdraw() + draw_label(i, fontface = 'bold', hjust = 0, size = 7),
      plot_grid(UMAP_grid[["UMAP4_split"]], UMAP_grid[["UMAPa_edit"]], ncol = 2),
      plot_grid(
        NULL,
        UMAP_grid[["UMAP4_legend"]],
        ncol = 2,
        rel_widths = c(0.1, 3)
      ),
      text,
      ncol = 1,
      nrow = 3,
      rel_heights = c(0.6, 4, 0.2, 0.3), rel_widths = c(1,1,1,1.4))
  print(UMAP5)
  print(paste0("EGFP+SNCA+ enrichment: ", paste(label2analyze, collapse = "")))
  rm(SeuInt3_PD)
}
```

```{r include=T}
saveRDS(SeuInt, paste0("Outputs/", date, "_SeuInt_1st_fil_int_mayt", UMIs.p.cell_lowthreshold, "lowt", UMIs.p.cell_highthreshold,"_lowt", percent.rb_threshold, "rb_SN.rds"))
saveRDS(SeuInt3, paste0("Outputs/", date, "_SeuInt3_1st_fil_int_mayt", UMIs.p.cell_lowthreshold, "lowt", UMIs.p.cell_highthreshold,"_lowt", percent.rb_threshold, "rb_SN.rds"))
saveRDS(colors_n, paste0("Outputs/", date, "_finalclusters_colors_SN.rds"))
gc()
```

## Conclusions:

In [this paper that evaluates QC metrics across multiple data sets](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9793662/) there is a clear difference in the fraction of mitochondrial reads between the brain (very low) and other tissues. Also Brain tissue appears with a higher number of active genes and a lower number of ribosomal reads. When it focus on the data of the SN, it appears that neurons have a higher number of active genes than glial cells (despite some specific neuronal clusters that appear to have the lowest number of active genes). Also, focusing on the bone marrow, T cells appears to have a higher percentage of ribosomal reads than other cell types.

Based on [here](https://nbisweden.github.io/excelerate-scRNAseq/session-qc/scRNAseq_QC_Asa_Bjorklund_2019.pdf) the number of detected genes correlates with cell size (in our case would be nuclei size). And based on [here](https://nba.uth.tmc.edu/neuroscience/m/s1/chapter08.html) "Neurons with long axons have a larger cell body and nucleus". All this may indicate that maybe clusters 0,4,6,7,8,12,18,9 may be neurons with large axons. IN this sense, maybe is not surpising that clusters 0,6,7,8,18 are enriched in EGFP expression. Because is more probable that neuronal cell types with long axons express DAT gene.

Maybe more more ribosomal percentage (including rRNA or ribosomal protein) may be related to the fact that there is RNA degradation, that is reflected in more free rRNA compared to non ribosomal RNA? I don;t understand it. There is something related to this [here](Possible that degradation of RNA leads to more templating of rRNA-fragments).
